### 2019-08-05 icse/overview.tex

\section{Overview}

% TODO
% \begin{figure}
%   \centering
%   \includegraphics[height=6em]{img/???}
%   \caption{}
%   \label{fig:array-literal}
% \end{figure}

\begin{itemize}
  \item Sytnax: automatically generate parser from ES-BNF
    \begin{itemize}
      \item ES-BNF features
      \item Packrat Parsers with 1-lookahead
      \item Example
    \end{itemize}
  \item Semantics: automatically convert abstract algorithms into Core programs
    \begin{itemize}
      \item Abstract algorithms
      \item Example
    \end{itemize}
\end{itemize}
\begin{itemize}
  \item Figure of overall structure
  \item Explanation of each module
\end{itemize}