### 2019-08-06 icse/main.tex

\documentclass[sigconf,review]{acmart}

\usepackage{kotex}
\usepackage[T1]{fontenc}
\usepackage{beramono}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{qtree}
\usepackage{ulem}
\usepackage{graphicx}
\usepackage{listings}

% removed ACM references
\settopmatter{printacmref=false}

% scala code style
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\lstdefinestyle{myScalastyle}{
  frame=tb,
  language=scala,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=fixed,
  basicstyle={\footnotesize\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  frame=single,
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3,
}
\lstdefinestyle{myCorestyle}{
  frame=tb,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=fixed,
  basicstyle={\footnotesize\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  frame=single,
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3,
}

% load macros
\input{macros}

\begin{document}

\title{Automatic Extraction of JavaScript Formal Semantics
from ECMAScript Specifications}

\author{Jihyeok Park}
\email{jhpark0223@kaist.ac.kr}
\affiliation{\institution{KAIST}}

\author{Jihee Park}
\email{j31d0@kaist.ac.kr}
\affiliation{\institution{KAIST}}

\author{Sukyoung Ryu}
\email{sryu.cs@kaist.ac.kr}
\affiliation{\institution{KAIST}}

\maketitle
\begin{abstract}
\end{abstract}

% TODO \keywords

\input{intro}
\input{overview}
\input{parser}
\input{compiler}
\input{framework}
\input{eval}
\input{related}
\input{conclusion}

\bibliographystyle{ACM-Reference-Format}
\bibliography{ref}

\end{document}