### 2019-08-07 icse/main.tex
Written intro

\documentclass[sigconf,review]{acmart}

\usepackage{kotex}
\usepackage[T1]{fontenc}
\usepackage{beramono}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{qtree}
\usepackage{ulem}
\usepackage{graphicx}
\usepackage{listings}

% removed ACM references
\settopmatter{printacmref=false}

% Scala code style
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\lstdefinestyle{myScalastyle}{
  frame=tb,
  language=scala,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=fixed,
  basicstyle={\footnotesize\ttfamily},
  numbers=none,
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  frame=single,
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3,
}

% 
\lstdefinestyle{ires}{
  frame=tb,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=fixed,
  basicstyle={\footnotesize\ttfamily},
  numbers=none,
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  frame=single,
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3,
}

% JavaScript code style
\lstdefinelanguage{JavaScript}{
  keywords={typeof, new, true, false, catch, function, return, null, catch, switch, var, if, in, while, do, else, case, break},
  keywordstyle=\color{blue}\bfseries,
  ndkeywords={class, export, boolean, throw, implements, import, this},
  ndkeywordstyle=\color{darkgray}\bfseries,
  identifierstyle=\color{black},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{purple}\ttfamily,
  stringstyle=\color{red}\ttfamily,
  morestring=[b]',
  morestring=[b]"
}

\lstdefinestyle{myJSstyle}{
  language=JavaScript,
  extendedchars=true,
  basicstyle=\footnotesize\ttfamily,
  showstringspaces=false,
  showspaces=false,
  numbers=none,
  tabsize=2,
  breaklines=true,
  showtabs=false,
  captionpos=b
}

% load macros
\input{macros}

\begin{document}

\title{Automatic Extraction of JavaScript Formal Semantics
from ECMAScript Specifications}

\author{Jihyeok Park}
\email{jhpark0223@kaist.ac.kr}
\affiliation{\institution{KAIST}}

\author{Jihee Park}
\email{j31d0@kaist.ac.kr}
\affiliation{\institution{KAIST}}

\author{Sukyoung Ryu}
\email{sryu.cs@kaist.ac.kr}
\affiliation{\institution{KAIST}}

\maketitle
\begin{abstract}
\end{abstract}

% TODO \keywords

\input{intro}
\input{overview}
\input{parser}
\input{compiler}
\input{framework}
\input{eval}
\input{related}
\input{conclusion}

\bibliographystyle{ACM-Reference-Format}
\bibliography{ref}

\end{document}