### 2019-08-19 icse/conclusion.tex
Written conclusion

\section{Conclusion}\label{sec:conclude}

We indiciate the problem that existing JavaScript semantics are not suitable for
annually updated cadence of ECMAScript specifications. We propose a tool \( \tool \),
JavaScript IR-based Semantics Extraction Toolchain to mechanize extracting syntax
and semantics from ECMAScript specifications. The most important modules are
the parser generator and the algorithm compiler. The parser generator generates
parsers from given \( \bnfes \) in lookahead parsers we proposed. the algorithm
compiler compiles given abstract algorithm written in structured natural language
into \( \ires \) functions. It is dependent on given compile rules and we provide
general compile rules for existing ECMAScript specifications. Moroever, we also
develop the rule assistant to diagnose root causes of not yet convertible algorithms.
It aids to modify compile rules for specification updates. We evaluate \( \tool \)
based on the next version of ECMAScript specification (2020) with the official conformance test
suite (test262). The extracted semantics passed all \inred{XXXXX} tests in test262
and we show that the tool is alo applicable into in-progress language features
by a case study. Moreover, we found \inred{four} real specification errors via \( \tool \)
and TC39 confirms all of them.