### 2019-08-24 icse20/conclusion.tex
Updated final numbers

\section{Conclusion}\label{sec:conclude}
The annual updates of the ECMAScript specification makes it difficult to
build program analysis or formal verification of JavaScript due to the
required human efforts in modeling a moving target.  In this paper,
we proposed a tool \( \tool \), which \textit{automatically} extracts the
syntax and semantics specified in the ECMAScript specification.  The tool
generates a parser from a language grammar written in \( \bnfes \), and
compiles abstract algorithms written in English to \( \ires \) functions.
The only parts that require manual efforts are compile rules and
language-specific global setting, which we provide for ECMAScript
as an example case study.  We also support a rule assistant to guide
how to convert new abstract algorithms.  We evaluated \( \tool \) by
applying it to ECMAScript 2020 with Test262.  The extracted semantics
passed 19,220 of 19,282 applicable tests in Test262 and we showed
that the tool is also applicable to an incomplete language proposal as
a case study.  The tool also detected six confirmed errors from the
specification and the proposal. The implementation of the tool is
publicly available as an open-source project.