### 2020-03-28 ase20/abstract.tex
[ASE] Revised abstract / intro

\begin{abstract}
JavaScript was initially designed for client-side programming in web browsers,
but its engine is now embedded in various kinds of host software.  Despite the
popularity, since the JavaScript semantics is complex especially due to its
dynamic nature, understanding and reasoning about JavaScript programs are
challenging tasks.  Thus, researchers have proposed several attempts to define
the formal semantics of JavaScript based on ECMAScript, the official JavaScript
specification.  However, the existing approaches are manual, labor-intensive,
and error-prone and all formal semantics targeted ECMAScript 5.1 (ES5.1, 2011)
or more former versions.  It is critical in understanding modern JavaScript
language features introduced by ECMAScript 6 (ES6, 2015).  Moreover, ECMAScript
has been annually released since 2015, which made already five updates after
ES5.1.

To alleviate the problem, we propose \( \tool \), a JavaScript IR-based
Semantics Extraction Toolchain.  It is the first tool that \textit{automatically
synthesizes} parsers and AST-IR translators directly from a given ECMAScript.
For syntax, we develop a parser generation technique with \textit{lookahead
parsing} for \( \bnfes \), a variant of the extended BNF used in ECMAScript.
For semantics, \( \tool \) synthesizes AST-IR translators using \textit{forward
compatible} rule-based compilation.  Each \textit{compile rule} describes how to
convert each step of abstract algorithms written in structured natural language
into \( \ires \), an Intermediate Representation that we designed for
ECMAScript.  For the five most recent ECMAScript versions, \( \tool \) succeeds
to synthesize parsers for all versions, and to compile \inred{XX.XX}\% of
algorithm steps in average.  We manually complete the core part of formal
semantics of the draft of the next ECMAScript (ES11, 2020) based on the
extracted semantics and it passed all \inred{XX,XXX} applicable tests.  Based on
the formal semantics, we found \inred{X} specification errors in ES11 and all of
them were confirmed by TC 39.  Furthermore, we also successfully applied \(
\tool \) to \inred{X} feature proposals in Stage 3 to check the forward
compatibility of \( \tool \).
\end{abstract}