### 2020-03-30 ase20/intro.tex
Revised Sec. 1

\section{Introduction}
JavaScript is one of the most widely used programming languages not only for
client-side but also for server-side programming~\cite{nodejs, meanjs} and even
for small embedded systems~\cite{espruino, tessel2}.  It is the top-ranked
language used in active GitHub repositories\footnote{https://githut.info/}, and
\#7 in the TIOBE Programming Community
index\footnote{https://www.tiobe.com/tiobe-index/}.  According to
W3Techs\footnote{https://w3techs.com/technologies/details/cp-javascript/all/all},
95.0\% of websites use JavaScript as their client-side programming language.

Despite its popularity, JavaScript developers often suffer from its intricate
semantics, which may cause unexpected behaviors.  For example, calling the
following JavaScript function seems to always return \( \code{false} \):
\begin{lstlisting}[style=myJSstyle]
        function f(x) { return x == !x; }
\end{lstlisting}
Unfortunately, it returns \( \code{true} \) when its argument
is an empty array \( \code{[]} \).  To correctly understand and reason about
such a complex behavior, the formal semantics of JavaScript is necessary.

\begin{figure}
  \centering
  \includegraphics[width=0.48\textwidth]{img/existing.pdf}
\vspace*{-2em}
  \caption{Existing approaches: Manually built parsers and AST-IR
  translators for JavaScript IR-based semantics}
  \label{fig:existing}
\vspace*{-1em}
\end{figure}

Researchers have defined various JavaScript formal
semantics~\cite{aplas08,lambdajs,kjs,javert} suitable for static
analysis~\cite{jsai,tajs,wala,safe} and formal verification~\cite{javert} by
referring to ECMAScript.  ECMAScript is the official specification that
describes the JavaScript syntax using a variant of the extended BNF (EBNF) notation,
and its semantics using abstract algorithms written in English in a clear and
structured manner.  \textit{IR-based semantics extraction} is a traditional way to
define the formal semantics of a language by building a compiler front-end
that takes programs and produces their Intermediate
Representations (IRs) to indirectly represent the semantics of given programs.
As illustrated in Figure~\ref{fig:existing}, the compiler front-end consists of a parser
that constructs Abstract Syntax Trees (ASTs) of given JavaScript programs, and
an AST-IR translator that converts ASTs to their own IRs.  This approach helps
researchers focus on IRs without worrying about diverse and enormous features of
JavaScript in developing new techniques for static analysis and formal
verification.

However, to the best of our knowledge, all existing approaches to JavaScript
IR-based semantics extraction \textit{manually} build parsers and translators.
Although it was a reasonable approach until ECMAScript 5.1 (ES5.1,
2011)~\cite{es5}, the manual approach is too tedious, labor-intensive, and
error-prone to deal with the large size of \textit{modern JavaScript} since ECMAScript 6
(ES6, 2015)~\cite{es6}. ES6 introduced numerous significantly new features such
as lexical binding via \( \code{let} \), the spread \( \code{...} \) operator,
classes, the \( \code{for-of} \) operator, the \( \code{async} \) functions, and
generators.  For example, consider KJS~\cite{kjs}.
KJS is one of formal semantics of ES5.1
and it is defined on top of \( \kframework \), which is a framework for
defining language semantics.  According to an author of KJS, it took
\textit{four months} to implement an AST-IR translator for 1,370 steps out of 2,932
steps in 368 abstract algorithms. However, the most recent version of ECMAScript
(ES10, 2019)~\cite{es10} has 1,731 abstract algorithms consisting of 10,068
steps. Thus, the manual approach does not seem to be scalable enough to build
an AST-IR translator for modern JavaScript
and indeed no formal semantics exists for ES6 to ES10.

Moreover, JavaScript syntax and semantics are annually updated.  Until
ES 5.1, JavaScript was a stable language because the specification was
rarely updated.  However, the Ecma Technical Committee 39 (TC39)~\cite{tc39}
decided to release the specification annually in late 2014.  After this official
announcement, several syntax features and roughly 1,000 to 3,000 steps of
abstract algorithms have been modified or newly added in the specification every year.
To handle these frequent and massive updates of ECMAScript,
the manual approach requires researchers to manually update parsers and AST-IR translators,
which incurs tremendous efforts.

To alleviate this problem, we propose a technique to \textit{automatically
synthesize} parsers and AST-IR translators directly from ECMAScript with
\textit{forward compatibility}. There are several technical challenges in
synthesizing parsers and translators. For syntax, ECMAScript utilizes its own
special variant of the EBNF with parametric non-terminals, conditional
alternatives, and various special terminal symbols.  Thus, no existing parser
generation technique is directly applicable for this variant.  Moreover,
JavaScript provides automatic semicolon insertion in its parsing algorithm with
several complex rules, not in a lexer. For semantics, abstract algorithms in
ECMAScript are written in English.  Besides, a general and forward compatible
representation of abstract algorithms is necessary to support future
versions of ECMAScript.

Our contribution is \( \tool \), a JavaScript IR-based Semantics Extraction
Toolchain:
\begin{itemize}[leftmargin=0.5cm]
  \item \textbf{\( \tool \) is the \textit{first tool that can automatically synthesize}
    parsers and AST-IR translators directly from a language specification,
    ECMAScript.}  For syntax, we formally introduce a variant of EBNF, \( \bnfes \),
    and propose a parser generation technique with
    lookahead parsing for \( \bnfes \), which supports automatic semicolon
    insertion. For semantics, we propose semi-automatic synthesis of AST-IR
    translators assisted by compile rules because of the
    difficulty in handling the semantics written in a natural language.  Compile rules describe how
    each step of abstract algorithms is converted into our intermediate
    representation \( \ires \) designed for ECMAScript. We evaluated \( \tool \)
    with the five most recent ECMAScript versions and the current draft of the next
    ECMAScript (ES6 to ES11).   \( \tool \) automatically generated
    parsers for all versions, and automatically compiled
    \inred{XX.X}\% of the steps in abstract algorithms on average.
  \item \textbf{ Using \( \tool \), we could define the \textit{first IR-based
    formal semantics of modern JavaScript}.}  The semantics targets the
    core of ES11, and \( \tool \) successfully synthesizes its
    parser and AST-IR translator for \inred{X,XXX} out of \inred{X,XXX}
    algorithm steps.  We manually completed the missing parts of the AST-IR
    translator (\inred{XXX} steps) in \inred{XX} days.  We evaluated the correctness
    of the formal semantics using Test262~\cite{test262}, the official
    conformance test suite, and the semantics passed all \inred{XX,XXX}
    applicable tests.  Moreover, we found \inred{X} specification errors in ES11
    thanks to the formal semantics.  They were all confirmed by TC39 and will be
    fixed in the next release.
  \item \textbf{\( \tool \) is also \textit{forward compatible} with new language features
    proposed for future ECMAScript specifications.}  TC39 manages proposals of
    new language features in an open source
    project\footnote{https://github.com/tc39/proposals}.  Out of five stages
    of proposals (from Stage 0 to Stage 4), proposals in Stage 4 (Finished)
    are ready for inclusiong in the next ECMAScript, and proposals in
    Stage 3 (Candidate) are in the last stage before finished.  We
    evaluated the forward compatibility of \( \tool \) by appling it
    to all \inred{XX} proposals in Stage 3. It successfully
    synthesized parsers for all proposals, and automatically
    compiled \inred{XX} out of \inred{XX} algorithm steps for all
    proposals.  After completion of the missing parts,
    each semantics passed all \inred{XX} tests.
\end{itemize}