### 2020-04-01 ase20/impl.tex

\section{Implementation}\label{sec:impl}
We describe the implementation detail of our open-source
tool, \( \tool \)\footnote{The URL of the tool is anonymized due to a
double-blind review process.}, which targets ES10~\cite{es10}.

%We describe how we implement an open-source \( \tool \)\footnote{
%The URL of the tool is anonymized due to a double-blind review process.},
%which targets ES10~\cite{es10}.

\subsection{Excluded Language Features}\label{sec:exclusion}
We decided not to support the following ES10 features:

\vspace*{-.5em}
\paragraph{Non-Strict Mode}
Since ES5.1, JavaScript introduced strict mode to
provide different semantics.  It
provides more stable semantics by preventing unsafe actions or
throwing originally ignored errors.  However, ECMAScript
states that ``Strict mode does not restrict or modify
any aspect of the ECMAScript semantics that must operate consistently
across multiple source text units.''  Thus, we decided to support only
strict mode.

\vspace*{-.5em}
\paragraph{Modules}
Moreover, we do not support modules that were introduced since ES6.
With the \( \code{import} \) keyword, a JavaScript
program can import other JavaScript components.  To support
modules in our automatically generated semantics, we should model the
functionality to communicate with file systems to access JavaScript
module files.  Because modules do not affect the overall JavaScript
semantics critically, we decided not to support modules.

\vspace*{-.5em}
\paragraph{Early Errors}
ECMAScript uses Early Error Rules to preclude invalid code
before evaluation. For example, an Early Error Rule checks whether
functions have duplicated parameters in strict mode.
Because they are written in different styles from abstract algorithms,
we do not support Early Error Rules.

\vspace*{-.5em}
\paragraph{Inessential Built-in Objects}
JavaScript provides various standard built-in objects but some of them
are not essential to core JavaScript but support extra functionalities such as
\( \code{RegExp} \) and \( \code{JSON} \).  Thus, we decided not to
support such libraries following the tradition of KJS~\cite{kjs},
which is a complete formal semantics of JavaScript.

\subsection{Compile Rules for ECMAScript}
\begin{table}[t]
  \centering
  \caption{Compile rules in ES10}
  \label{table:rules}
\vspace*{-1em}
  \[
    \begin{array}{c?r|r|r|r|r|r|r}
      \text{Name}
      & \multicolumn{1}{c|}{\code{Stmt}}
      & \multicolumn{1}{c|}{\code{Expr}}
      & \multicolumn{1}{c|}{\code{Cond}}
      & \multicolumn{1}{c|}{\code{Value}}
      & \multicolumn{1}{c|}{\code{Ty}}
      & \multicolumn{1}{c|}{\code{Ref}}
      & \multicolumn{1}{c}{\code{SecNo}}\\\hline
      \text{\# Rules}
      & 17
      & 16
      & 8
      & 11
      & 33
      & 7
      & 21
    \end{array}
  \]
\vspace*{-2em}
\end{table}

As we discussed in Section~\ref{sec:compiler}, \textsf{Algorithm Compiler}
requires compile rules to compile given abstract algorithms to \( \ires \)
functions.  In order to cover the entire ES10,
we defined compile rules for only seven kinds as summarized in
Table~\ref{table:rules}.  The compile rule for statements, \( \code{Stmt} \),
contains 17 compile rules and generates \( \ires \) instructions.
The \( \code{Expr} \), \( \code{Cond} \), and \( \code{Value} \) compile
rules denote expressions, conditions, and values, respectively.  While
all of them generate \( \ires \) expressions, they represent different
contexts in ECMAScript.  For example,
\( \code{Expr} \) represents a context where any expression can appear,
\( \code{Cond} \) denotes a context where any boolean-valued expression
can appear, and \( \code{Value} \) represents a context where a fully
evaluated value can appear.  The \( \code{Ty} \) compile rule denotes
type names and generates string primitives in \( \ires \).  The \( \code{Ref} \)
represents references such as identifier lookup and member accesses of
map-like structures.  It generates \( \ires \) reference expressions.
Finally, the \( \code{SecNo} \) compile rule denotes section numbers
like \( \code{9.5.12} \), and it converts such numbers to
corresponding name strings like \( \code{"ProxyExoticObject.Call"} \).
% \begin{lstlisting}[style=myScalastyle]
% lazy val SecNo =
%   "9.5.12" ^^ { case x => "ProxyExoticObject.Call" } |||
%   ...
% \end{lstlisting}

\subsection{Global Setting}
We manually implemented \textsf{Global Setting} for ES10,
which has two parts: ECMAScript data types and built-in objects.

\vspace*{-.5em}
\paragraph{ECMAScript Data Types}
In ES10, Section 6 \textsf{ECMAScript Data Types and Values}
describes data types used in the specification.  It describes not only
data types explicitly exposed for JavaScript evaluation, but also all
the internal data types used in the specification.  For example, the
Environment Records are record types from identifiers to bindings for
values.  They are not explicitly accessible in JavaScript programs
but defined and used internally in the specification.

Each ECMAScript data type has several fields to store their relevant values.
We should implement such structures in order to correctly evaluate JavaScript
programs, but fields of ECMAScript data types could be abstract algorithms.
We call them \textit{method-like abstract algorithms} because they
implicitly get their receiver objects as arguments at callsites.
To mimic such implicit behavior, we added a special variable
\( \code{this} \) as the first parameter of a method-like abstract algorithm,
and passed a receiver object at its callsite by modifying
\textsf{Algorithm Compiler}.  For example, an Environment Record has
the \textsf{DeleteBinding (N)} abstract algorithm as one of its fields.
Then, it gets two parameters, the special parameter \( \code{this} \)
and normal one \( \code{N} \).  Thus, the following algorithm step:
\begin{center}
  \includegraphics[width=0.2\textwidth]{img/method-like-example.png}
\end{center}
is compiled into the following \( \ires \) instruction:
\begin{lstlisting}[style=ires]
return DclRec.DeleteBinding(DclRec, N)
\end{lstlisting}

\vspace*{-1em}
\paragraph{Built-in Objects}
In ECMAScript, built-in objects are pre-defined functions with several
built-in functions.  For example, \( \code{Array} \)
is the constructor of array objects, and its prototype
\( \code{Array.prototype} \) has built-in functions for array objects.
For instance, \( \code{[1,2,3].flat()} \) calls the
\( \code{Array.prototype.flat} \) built-in function with the array
\( \code{[1,2,3]} \).  Because built-in functions are also abstract
algorithms, each of them is automatically converted to an \( \ires \)
function.  However, the structures of built-in objects should be
manually implemented.  Thus, we implemented built-in objects in Scala
and connected their properties with the extracted \( \ires \) functions.
%
Some built-in objects that are explicitly referenced in abstract
algorithms are intrinsic objects, which have their own
aliased names summarized in Table 7 of Section 6.1.7.4
\textsf{Well-Known Intrinsic Objects}.  We extracted the alias
into \textsf{Global Setting} to utilize it during evaluation.