### 2020-04-03 ase20/overview.tex
Revised Sec. 2

\section{Overview}
\begin{figure}
  \centering
  \includegraphics[width=0.48\textwidth]{img/overview.pdf}
  \caption{Overall structure of \( \tool \): Automatically synthesized {\sf
    JavaScript Parser} and {\sf AST-\( \ires \) Translator} for JavaScript
    IR-based semantics}
  \label{fig:overview}
\vspace*{-1em}
\end{figure}

In this section, we introduce the overall structure of \( \tool \)
depicted in Figure~\ref{fig:overview}.
Compared to the existing approaches shown in
Figure~\ref{fig:existing}, \( \tool \) automatically synthesizes
{\sf JavaScript Parser} and {\sf AST-\( \ires \) Translator} directly from
ECMAScript.

The motivation of automatically extracting syntax and semantics of JavaScript
from ECMAScript is twofold: 1) ECMAScript is written in a well-organized
style, and 2) it is available as HTML files using various HTML tags.
To easily manipulate ECMAScript written in HTML, we developed
{\sf Spec Extractor} in JavaScript to extract the syntax and semantics in JSON format.

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/arrayliteral-syntax.pdf}
    \caption{\textit{ArrayLiteral} production in ES10}
    \label{fig:array-literal-es}
  \end{subfigure}
  \begin{subfigure}[t]{0.48\textwidth}
    \begin{lstlisting}[style=myScalastyle]
val ArrayLiteral: List[Boolean] => LAParser[T] = memo {
  case List(Yield, Await) =>
  "[" ~ opt(Elision) ~ "]"             ^^ ArrayLiteral0 |
  "[" ~ ElementList(Yield,Await) ~ "]" ^^ ArrayLiteral1 |
  "[" ~ ElementList(Yield,Await) ~ ","
      ~ opt(Elision)~ "]"              ^^ ArrayLiteral2
}
    \end{lstlisting}
    \vspace*{-1em}
    \caption{Generated parser for the \textit{ArrayLiteral} production}
    \label{fig:array-literal-parser}
  \end{subfigure}
  \vspace*{-1em}
  \caption{\textit{ArrayLiteral} production in ES10 and its parser}
  \label{fig:array-literal}
\end{figure}

\smallskip

\textbf{Syntax.} ECMAScript provides the lexical and syntactic grammars in
Appendix A using a variant of EBNF for ECMAScript.  We dub it \( \bnfes \) and
formally define it in Section~\ref{sec:parser}.  Our {\sf Spec Extractor} reads
the grammars written in \( \bnfes \) and converts them into JSON files.
For example, Figure~\ref{fig:array-literal}(a) shows the \textit{ArrayLiteral}
production in ES10.  It takes two boolean parameters \textsf{Yield} and
\textsf{Await} and has three alternatives.  The first alternative
consists of three symbols: two terminal symbols \( \code{[} \) and
\( \code{]} \), and one non-terminal symbol \textit{Elision}$_{\mbox{\footnotesize opt}}$.
The {\small opt} subscript denotes that it is optional.  In the second
and third alternatives, \textit{ElementList}$_{\mbox{\footnotesize\sf [?Yield, ?Await]}}$
denotes a parametric non-terminal symbol \textit{ElementList}
with the parameters \textsf{Yield} and \textsf{Await} of
\textit{ArrayLiteral} as its two arguments. The prefix \textsf{\small ?}
of a symbol denotes that the symbol is passed as an argument.

To generate {\sf JavaScript Parser} from a given \( \bnfes \) grammar, we
construct \textsf{Parser Generator} in Scala.  It synthesizes a
JavaScript parser according to the given \( \bnfes \), and the
generated parser is defined with Scala parser combinators~\cite{scala-parser-combinators}.
Moreover, in order to parse \( \bnfes \) grammars correctly and
efficiently, we propose \textit{lookahead parsers}, which keep track
of lookaheads, sets of possible next tokens.  With lookahead parsing,
generated parsers now have one-to-one mapping to their corresponding
grammar productions, improving readability.
For example, Figure~\ref{fig:array-literal}(b) shows the generated
parser for the \textit{ArrayLiteral} production in Figure~\ref{fig:array-literal}(a).
Each parser has the \( \code{List[Boolean] => LAParser[T]} \) type
because each production in \( \bnfes \) is parametric with boolean values.
The \( \code{memo} \) is a memoization function for pairs of boolean
parameters and resulting parsers for performance optimization.
The value \( \code{ArrayLiteral} \) corresponds to the
\textit{ArrayLiteral} production.  In the parser, each string literal
such as \( \code{"["} \) or \( \code{"]"} \) denotes a parser for a
terminal symbol.  The \( \code{opt} \) helper function creates
optional parsers.  The parametric non-terminal \textit{ElementList}
with arguments \textsf{Yield} and \textsf{Await} is represented as a
function call \( \code{ElementList(Yield, Await)} \).
The \( \code{\textasciitilde} \) operator combines two parsers
and the \( \code{\^{}\^{}} \) operator describes how to construct ASTs.
When the left-hand side of \( \code{\^{}\^{}} \) is matched, its
right-hand side shows a corresponding AST constructor, where the name
of each constructor has a number denoting the order among alternatives.
For example, the \( \code{ArrayLiteral0} \) constructor corresponds to
the first alternative of the \textit{ArrayLiteral} production.

\begin{figure}
  \centering
  \begin{subfigure}{0.43\textwidth}
    \includegraphics[width=\textwidth]{img/arrayliteral-eval.pdf}
    \subcaption{\textbf{Evaluation} abstract algorithm for the third alternative}
    \label{fig:array-literal-eval-es}
  \end{subfigure}
  \qquad
  \begin{subfigure}{0.46\textwidth}
    \begin{lstlisting}[style=ires]
ArrayLiteral[2].Evaluation (ElementList, Elision) => {
  let array = ! (ArrayCreate 0)
  let len = (ElementList.ArrayAccumulation array 0)
  ? len
  if (= Elision absent) let padding = 0
  else let padding = Elision.ElisionWidth
  (Set array "length" (ToUint32 (+ padding len)) false)
  return array
}
    \end{lstlisting}
    \vspace*{-1em}
    \subcaption{The generated \( \ires \) function}
    \label{fig:array-literal-eval-ires}
  \end{subfigure}
  \vspace{-1em}
  \caption{\textbf{Evaluation} abstract algorithm for the third alternative of
    \textit{ArrayLiteral} in ES10 and its generated \( \ires \)
    function}
  \label{fig:array-literal-eval}
  \vspace{-1em}
\end{figure}

\textbf{Semantics.}
ECMAScript describes the language semantics as abstract algorithms in English.
While they are written in a natural language, the writing style is
well-organized with ordered steps and tagged tokens.  {\sf Spec Extractor}
reads abstract algorithms with HTML tags and converts them into JSON files.
For example, Figure~\ref{fig:array-literal-eval}(a) presents the
\textbf{Evaluation} abstract algorithm of the third alternative of the
\textit{ArrayLiteral} production in ES10, and it has seven steps.  In the HTML
files describing the abstract algorithms, each non-terminal symbol
(e.g. \textit{ElementList}), local variable (e.g. \textit{array}), code (e.g. \(
\code{"length"} \)), or value (e.g. \textbf{false}) has the \( \code{<nt>} \),
\( \code{<var>} \), \( \code{<code>} \), or \( \code{<emu-val>} \) tag,
respectively.

To translate such abstract algorithms into representations
suitable for manipulation, we define \( \ires \), a specialized
intermediate representation for ECMAScript.  Then,
we develop \textsf{Algorithm Compiler} in Scala using Scala parser
combinators again to convert given abstract algorithms to \( \ires \)
functions.  It also takes \textsf{Compile Rules} as another input,
which has two parts: parsing rules and conversion rules.  They are
\textit{manually} specified for ECMAScript; we explain them in
detail in Section~\ref{sec:compiler}.  Thus, each abstract algorithm is
converted into a function written in \( \ires \) via \textsf{Algorithm Compiler}.
For example, Figure~\ref{fig:array-literal-eval}(b) presents the generated
\( \ires \) function for the \textbf{Evaluation} abstract algorithm shown in
Figure~\ref{fig:array-literal-eval}(a).
The \( \code{ArrayLiteral[2].Evaluation} \) function takes two parameters for
two non-terminal symbols: \( \code{ElementList} \) and \( \code{Elision} \).
The parameter \( \code{Elision} \) has a special value \( \code{absent} \) when
the non-terminal symbol \textit{Elision}$_{\mbox{\footnotesize opt}}$ is not
present.  Thus, we convert the condition in step 4,
``if \textit{Elision} is not present,'' into the equality check with \(
\code{absent} \): \( \code{if (= Elision absent)} \).  Codes are represented as
string values and values are represented as corresponding \( \ires \) values.
For instance, the code \texttt{\small "length"} and the value \textbf{false} are
converted into the string value \( \code{"length"} \) and the boolean value
\( \code{false} \), respectively.

Finally, \( \tool \) constructs {\sf AST-\( \ires \) Translator} with the given
\( \ires \) functions and \textit{manually} specified {\sf Global Setting},
which has minor but necessary information to evaluate JavaScript programs
described in ECMAScript such as the structure of the standard
built-in objects and ECMAScript data types. Putting them all together, we can
translate a given JavaScript program into \( \ires \) via generated {\sf
JavaScript Parser} and {\sf AST-\( \ires \) Translator} by \( \tool \).  Even
though \( \tool \) is not fully automatic because of {\sf Compile Rules} and
{\sf Global Setting}, it could dramatically reduce the efforts to building
parsers and translators from scratch.

In the remainder of this paper, we explain the details of how to
automatically generate parsers (Section~\ref{sec:parser}) and how to compile
abstract algorithms (Section~\ref{sec:compiler}).  After describing implementation
details (Section~\ref{sec:impl}), we evaluate \( \tool \)
(Section~\ref{sec:eval}), discuss related work (Section~\ref{sec:related}), and
conclude (Section~\ref{sec:conclude}).