### 2020-04-30 ase20/eval.tex
Till Sec. 5.2

\section{Evaluation}\label{sec:eval}

We developed \( \tool \) as an open-source tool\footnote{The URL of the tool is
anonymized due to a double-blind review process.}, and evaluated the tool
based on the following research questions:
\begin{itemize}[leftmargin=0.5cm]
\item RQ1. \textbf{Coverage:} How much percentage of 
the syntax and semantics does \( \tool \) automatically extract from ES7 to ES10?
\item RQ2. \textbf{Correctness:} Does \( \tool \) extract an IR-based formal
semantics from ECMAScript correctly?
\item RQ3. \textbf{Forward Compatibility:} Is \( \tool \) applicable to language
features ready for inclusion in the next ECMAScript (ES11)?
\end{itemize}

\subsection{Coverage}

\begin{table}[t]
  \caption{Syntax coverage: Number of productions in each specification
and in each update between adjacent versions,
from \textit{all} of which \( \tool \) automatically generated parsers}
  \label{fig:syntax-all-version}
\vspace*{-1em}
\small
  \[
    \begin{array}{c?r|r|r|r?r}
      \textbf{Version}
      & \multicolumn{1}{c|}{\textbf{ES7}}
      & \multicolumn{1}{c|}{\textbf{ES8}}
      & \multicolumn{1}{c|}{\textbf{ES9}}
      & \multicolumn{1}{c?}{\textbf{ES10}}
      & \multicolumn{1}{c}{\textbf{Average}}\\\toprule
      \text{\# Lexical productions}
      & \inred{\text{78}}
      & \inred{\text{78}}
      & \inred{\text{78}}
      & \inred{\text{81}}
      & \inred{\text{78.75}}\\\hline
      \text{\# Syntactic productions}
      & \inred{\text{157}}
      & \inred{\text{167}}
      & \inred{\text{167}}
      & \inred{\text{174}}
      & \inred{\text{166.25}}\\
    \end{array}
  \]
  \[
    \begin{array}{c?r|r|r?r}
      \textbf{Old version}
      & \multicolumn{1}{c|}{\textbf{ES7}}
      & \multicolumn{1}{c|}{\textbf{ES8}}
      & \multicolumn{1}{c?}{\textbf{ES9}}
      & \multirow{2}{*}{\textbf{Average}}\\\cline{1-4}
      \textbf{New version}
      & \multicolumn{1}{c|}{\textbf{ES8}}
      & \multicolumn{1}{c|}{\textbf{ES9}}
      & \multicolumn{1}{c?}{\textbf{ES10}}
      & \\\toprule
      \Delta \; \text{\# Lexical productions}
      & \inred{\text{3}}
      & \inred{\text{5}}
      & \inred{\text{6}}
      & \inred{\text{4.67}}\\\hline
      \Delta \; \text{\# Syntactic productions}
      & \inred{\text{140}}
      & \inred{\text{15}}
      & \inred{\text{8}}
      & \inred{\text{54.33}}\\
    \end{array}
  \]
\end{table}

We evaluated the coverage of \( \tool \) in two respects: syntax
and semantics.  For syntax, we measured how many grammar
productions in specifications \( \tool \) automatically generated
parsers from, and for semantics, we measured how many abstract
algorithm steps in specifications it automatically generated the
JavaScript semantics from.  Because \( \tool \) utilizes common
patterns in the converged writing style since ES7 as we discussed in
Section~\ref{sec:overview}, we evaluated its coverage using the most
recent four versions of ECMAScript, ES7 to ES10.  We measured the
numbers for each ECMAScript version and for each update between
adjacent versions.

For syntax, \( \tool \) automatically generated parsers for \textit{all} the
lexical and syntactic productions.  As Table~\ref{fig:syntax-all-version} shows,
the average numbers of lexical and syntactic productions are \inred{78.75} and
\inred{166.25}, respectively.  Also, the average numbers of annually updated
lexical and syntactic productions between adjacent versions are \inred{4.67} and
\inred{54.33}, respectively.

\begin{figure}[t]
  \centering
  \begin{subfigure}{0.48\textwidth}
    \includegraphics[width=\textwidth]{img/sem-all.pdf}
    \caption{For each ECMAScript version from ES7 to ES10}
\vspace*{.5em}
    \label{fig:semantics-all-version}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.48\textwidth}
    \includegraphics[width=\textwidth]{img/sem-update.pdf}
    \caption{For each update between adjacent versions}
    \label{fig:semantics-all-version-update}
  \end{subfigure}
\vspace*{-.5em}
\caption{Semantics coverage: Number of algorithm steps in specifications,
from which \( \tool \) generated the semantics}
  \label{fig:all-version}
\vspace*{-1em}
\end{figure}

For semantics, Figure~\ref{fig:all-version} shows that \( \tool \) automatically
compiled algorithm steps to corresponding \( \ires \) instructions with
the success rate of \inred{91.38\%} on average for each ECMAScript version
from ES7 to ES10, and \inred{92.08\%} for each update between adjacent
versions.  ECMAScript abstract algorithms describe not only
core language semantics but also built-in libraries with various helper functions.
Note that built-in libraries are written in more diverse styles than
core language semantics due to their own specific functionalities.
Therefore, built-in libraries have lower success rates
(\inred{86.74\%} for specifications and \inred{88.24\%} for updates) than core
language semantics (\inred{95.14\%} for specifications and \inred{96.41\%} for
updates).

Because \( \tool \) automatically extracts the syntax
and semantics from specifications and updates with high coverage rates,
it reduces efforts not only in developing JavaScript tools from scratch
from specifications but also in evolving existing tools for updates.

\subsection{Correctness}\label{sec:check-with-tests}
\begin{table*}[t]
  \centering
  \caption{Specification errors in ES10 and the BigInt proposal ready for
  inclusion in ES11}
  \label{table:spec-errors}
  \vspace*{-.5em}
  \small
  \begin{tabular}{@{}c?c|l|@{}c@{}|c|c|r|@{}r@{}}
\multicolumn{1}{@{}c?}{\bf Name} &
\multicolumn{1}{c}{\bf Feature} &
\multicolumn{1}{c}{\bf Description} &
\multicolumn{1}{@{}c@{}}{\bf Known} &
\multicolumn{1}{c}{\bf Created} &
\multicolumn{1}{c}{\bf Resolved} &
\multicolumn{1}{c}{\bf Existed} &
\multicolumn{1}{@{}c@{}}{\bf \# Fails} \\\toprule
&&&&&&&\\[-1.5em]
    ES10-1 &
    Iteration &
    \makecell[l]{Missing the \code{async-iterate} case in the assertion of\\
 {\bf ForIn/OfHeadEvaluation}} &
    \textsf{X} &
    2018-02-16 &
    2020-03-25 &
    768 days &
    1,120 \\\hline

    ES10-2 &
    Condition &
    \makecell[l]{Ambiguous grammar production for the dangling \code{else} problem in\\
    {\it IfStatement}} &
    \textsf{X} &
    2015-06-01 &
    TBD &
    \multicolumn{1}{c|}{TBD} &
    1 \\\hline

    ES10-3 &
    String &
    \makecell[l]{Wrong use of the \code{=} operator in {\bf
    StringGetOwnProperty}} &
    \textsf{X} &
    2015-06-01 &
    TBD &
    \multicolumn{1}{c|}{TBD} &
    \inred{XX} \\\hline

 ES10-4 &
 Completion &
 Unhandling abrupt completion in {\bf Abstract Equality Comparison} &
 \textsf{X} &
\inred{0000-00-00}&
 2020-04-28 &
 \inred{XX} days &
 \inred{XX}
 \\\hline

    ES10-5 &
    Completion &
    \makecell[l]{Unhandling abrupt completion in {\bf Evaluation} of {\it
    EqualityExpression}} &
    \textsf{O} &
    2015-06-01 &
    2019-05-02 &
    1,065 days &
    1 \\\hline

    ES10-6 &
    Await &
    \makecell[l]{Passing a value of wrong type to the second parameter of\\ {\bf PromiseResolve}} &
    \textsf{O} &
    2019-02-27 &
    2019-04-13 &
    45 days &
    1,100 \\\hline

    ES10-7 &
    Function &
    \makecell[l]{No semantics of {\bf IsFunctionDefinition} for
    \code{function(...)\{...\}}} &
    \textsf{O} &
    2015-10-30 &
    2020-01-18 &
    1,541 days &
    \inred{XX} \\\hline

    ES10-8 &
    Function &
    \makecell[l]{No semantics of {\bf ExpectedArgumentCount} for the base case
    of\\ {\it FormalParameters}} &
    \textsf{O} &
    2016-11-02 &
    2020-02-20 &
    1,205 days &
    \inred{XX} \\\hline

    ES10-9 &
    Iteration &
    \makecell[l]{Two semantics of {\bf VarScopedDeclarations} for\\ \code{for
    await(var x of e)\{...\}}} &
    \textsf{O} &
    2018-02-16 &
    2019-10-11 &
    602 days &
    \inred{XX} \\\hline\hline

    BigInt-1 &
    Expression &
    \makecell[l]{Using wrong variable \code{oldvalue} instead of \\
    \code{oldValue} in {\bf Evaluation} of {\it UpdateExpression}.} &
    \textsf{X} &
    2019-09-27 &
    2020-04-23 &
    209 days &
    \inred{XX} \\\hline

    BigInt-2 &
    Number &
    \makecell[l]{Using {\bf ToInt32} instead of {\bf ToUint32} \\ in {\bf
    Number::unsignedRightShift}.} &
    \textsf{X} &
    2019-09-27 &
    2020-04-23 &
    209 days &
    \inred{XX} \\\hline

    BigInt-3 &
    Number &
    \makecell[l]{Using {\bf ToUint32} instead of {\bf ToInt32} in {\bf
    NumberBitwiseOp}.} &
    \textsf{X} &
    2019-09-27 &
    2020-04-23 &
    209 days &
    \inred{XX} \\\hline

    BigInt-4 &
    Number &
    \makecell[l]{Not handling BigInt values in {\bf Number} constructor.} &
    \textsf{O} &
    2019-09-27 &
    2019-11-19 &
    53 days &
    \inred{XX}
  \end{tabular}
\end{table*}

\begin{table}[t]
  \centering
  \caption{Test results for Test262}
  \label{table:test262}
  \vspace*{-.5em}
  \small
  \begin{tabular}{lr}\toprule
    \belowrulesepcolor{gainsboro}
    \rowcolor{gainsboro} \textbf{All Test262 Tests} & \inred{\textbf{XX,XXX}}\\
    \aboverulesepcolor{gainsboro}\midrule
    Annexes & \inred{XXX}\\\hdashline
    Internationalization & \inred{XXX}\\\hdashline
    In-progress features & \inred{X,XXX}\\\midrule
    \belowrulesepcolor{gainsboro}
    \rowcolor{gainsboro} \textbf{ES10 Tests} & \inred{\textbf{XX,XXX}}\\
    \aboverulesepcolor{gainsboro}\midrule
    Non-strict mode & \inred{X,XXX}\\\hdashline
    Modules & \inred{X,XXX} \\\hdashline
    Early errors before actual execution & \inred{X,XXX} \\\hdashline
    Inessential built-in objects & \inred{X,XXX} \\\midrule
    \belowrulesepcolor{gainsboro}
    \rowcolor{gainsboro} \textbf{Applicable Tests} & \inred{\textbf{XX,XXX}}\\
    \aboverulesepcolor{gainsboro}\midrule
    Passed tests & \inred{XX,XXX} \\\hdashline
    Failed tests & \inred{XXX} \\\bottomrule
  \end{tabular}
  \vspace*{-1em}
\end{table}

To evaluate the correctness of \( \tool \), we checked whether the
extracted semantics from the latest ECMAScript (ES10) passes Test262
as of February 28, 2019, when ES10 was branched out.
To focus on the core language semantics of JavaScript, we completed
only necessary parts missing from the extracted AST-IR translator.
As Figure~\ref{fig:all-version}(a) shows, for the abstract algorithms
in ES10, \inred{8,901} steps out of \inred{9,674} steps are
automatically compiled by \textsf{Algorithm Compiler}.
Among the remaining \inred{773} steps, we manually implemented all
missing steps for the core language semantics (\inred{221} steps) and
the essential parts of the built-in libraries (\inred{XX} steps out of
\inred{552}).  We also manually implemented \textsf{Global Setting}
as described in Section~\ref{sec:compiler-impl} for the core language features.
Note that we do not support minor language features such as the
non-strict mode, modules, early errors before actual execution, and
inessential built-in objects.
Among \inred{XX,XXX} tests in Test262, we filtered out \inred{XX,XXX}
tests as summarized in Table~\ref{table:test262}.  To focus on ES10,
we excluded \inred{X,XXX} tests for annexes, internationalization, and
in-progress features.  We also filted out \inred{X,XXX} tests that use
minor language features.
Finally, for \inred{XX,XXX} applicable tests, the extracted semantics
failed for \inred{XXX} tests.

We investigated the failed tests and found out that they failed due to
specification errors in ES10.  Using the failed tests, we
discovered nine errors: ES10-1 to ES10-9 in
Table~\ref{table:spec-errors}.  Among them, five errors (ES10-5 to
ES10-9) were previously reported and fixed in the current draft of the
next ECMAScript, and the remaining four errors (ES10-1 to ES10-4) were
never reported before.  All four errors were confirmed by TC39, and
will be fixed in the next ECMAScript, ES11.

The specification error ES10-1 is due to a wrong assertion.
While ES9 introduced the \code{for await} iteration statement with a
new {\it iterationKind} tag, \code{async-iterate}, the
\code{ForIn/OfHeadEvaluation} algorithm missed the
\code{async-iterate} case in an assertion, which caused 1,120
tests failed.  We reported the error and proposed a specification fix
to include  the \code{async-iterate} case, and TC39 accepted it on
March 25, 2020.  Because the error was created on February 16, 2018,
it existed for 768 days.

ES10-2 comes from the well-known dangling \code{else} problem
introduced in ALGOL 60~\cite{dangling-else}.  ES10 describes how to
parse it in prose: the \code{else} statement should be associated with
the nearest \code{if} statement.  Because it is written in prose
rather than in the ES10 grammar productions, it caused one test
failed.  We proposed a fix to revise the ambiguous grammar production,
and TC39 confirmed it on April 23, 2020.

ES10-3 is due to a misuse of the \code{=} operator for numbers.
In abstract algorithms, ``{\it x} = {\it y}'' denotes equality testing
for double-precision 64-bit binary format IEEE 754-2008 values;
thus, \mbox{``+0 = -0''} evaluates to true.  However, to check whether
{\it index} is exactly the same with -0,
{\bf StringGetOwnProperty} used \mbox{``{\it index} = -0''},
which is true even when {\it index} is +0.
It caused \inred{XX} tests failed.  We proposed a fix, which is
confirmed on April 22, 2020.

ES10-4 and ES10-5 happened because ES10 did not handle
abrupt completion from function calls.  Our proposed fix to ES10-4 was
accepted on April 28, 2020, and ES10-5 was resolved on May 2, 2019
after existing for 1,065 days.

ES10-6 is due to incorrect uses of an abstract algorithm.
While {\bf PromiseResolve(C, x)} expects a JavaScript object for its
second argument, ES10 passed a list of values rather than an object in
three invocations of {\bf PromiseResolve}.  The wrong invocations were
introduced on February 27, 2019 and caused \inred{XX} tests failed.
They were fixed on April 13, 2019 after existing 45 days.

ES10-7 and ES10-8 happened because ES10 missed semantics for some
cases.  They both existed for more than 1,200 days.

ES10-9 is due to multiple semantics.  Because only one semantics is
intended, \inred{XX} tests fail when a wrong semantics is chosen.

After resolving the nine specification errors in ES10, we extracted a
semantics from the revised specification.  The extracted semantics
from the revised ES10 passed all \inred{XX,XXX} applicable tests in Test262,
which shows that \( \tool \) extracts an IR-based formal semantics
from ECMAScript correctly.  In addition, the evaluation witnesses that
\( \tool \) can detect specification errors effectively.
We could detect not only five previously-known errors
but also four new errors.  We believe that \( \tool \) bridges gaps
between ECMAScript written in a natural language and executable tests
in Test262.

\subsection{Forward Compatibility}

We show the forward compatibility of \( \tool \) by applying it to the proposals
that are ready for inclusion in the next ECMAScript (ES11, 2020). Because
ECMAScript administered as an open-source project, various proposals for new
language features are available with their own specification changes and tests.
A separate repository~\cite{proposals} maintains such proposals with six
different stages: Stage 0 to Stage 3, Finished, and Inactive.  Each proposal
starts with Stage 0, and the TC39 committee regularly examines Stage 3 proposals
to decide their next stages.  If a proposal is confirmed, the committee changes
its stage to Finished and integrates it into the next version of ECMAScript.
Otherwise, the committee changes its stage to Inactive.  Among them, we applied
\( \tool \) to all Finished proposals decided to be included in ES11.

\begin{table}[t]
  \centering
  \caption{Evaluation result of \( \tool \) for Finished stage proposals that
  will be included in ES11.}
  \label{table:spec-prop-result}
  \vspace*{-0.5em}
  \small
  \begin{tabular}{c?r|r?r?r?r}
    \hline
    \multicolumn{1}{c?}{\multirow{2}{*}{\bf Name}} &
    \multicolumn{2}{c?}{\bf \( \Delta \) \# Prod.} &
    \multicolumn{1}{c?}{\multirow{2}{*}{\bf \( \Delta \) \# Steps}} &
    \multicolumn{1}{c?}{\multirow{2}{*}{\bf \( \Delta \) \# Tests}} &
    \multicolumn{1}{c}{\multirow{2}{*}{\bf \# Tests}} \\\cline{2-3}

    &
    \multicolumn{1}{c|}{\bf Lex.} &
    \multicolumn{1}{c?}{\bf Syn.} &
    &\\\hline\hline

    \code{matchAll} of \code{String} &
    0 &
    0 &
    9/9 &
    5/5 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    \code{import()} &
    0 &
    2 &
    22/23 &
    0/0 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    \code{BigInt} &
    4 &
    0 &
    258/323 &
    \inred{207}/207 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    \code{Promise.allSettled} &
    0 &
    0 &
    79/85 &
    50/50 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    \code{globalThis} &
    0 &
    0 &
    1/1 &
    1/1 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    \code{for-in} mechanics &
    0 &
    0 &
    36/37 &
    0/0 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    Optional Chaining &
    3 &
    3 &
    72/74 &
    19/19 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    Nullish Coalescing Op. &
    1 &
    4 &
    10/10 &
    21/21 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    \code{import.meta} &
    0 &
    2 &
    15/15 &
    0/0 &
    \inred{XXXXX}/\inred{XXXXX}\\\hline

    {\bf Total} &
    8 &
    11 &
    502/577\\\cline{1-4}
  \end{tabular}
\end{table}

Table~\ref{table:spec-prop-result} shows the result of applying \( \tool \) to 9
Finished proposals for ES11.  For all proposals, \inred{XX} lexical and
\inred{XX} syntactic productions are modified and \( \tool \) successfully
updated the changes to the synthesized JavaScript parser.  The updated parser
parses all applicable tests for all proposals.  For abstract algorithms,
\inred{XXX} steps out of \inred{XXX} steps are automatically converted to the
corresponding \( \ires \) instructions by the algorithm compiler without
changing of the compile rules.  Thus, our tool has \inred{XX.XX\%} success rate
on average for future proposals.  Similar with
Section~\ref{sec:check-with-tests}, we manually implemented missing part of the
AST-IR translator for each Finished proposal and checked each extracted
semantics with official tests.  Each of them passed all applicable tests except
the semantics for \code{BigInt} proposal.  It fails \inred{XX} tests out of 207
applicable tests for the proposal and \inred{XXXX} tests out of \inred{XXXXX}
applicable tests for ES10.

Based on the failed tests, we found three undiscovered specification errors
(BigInt-1, BigInt-2, and BigInt-3), and one reported error (BigInt-4) in the
BigInt proposal as described in Table~\ref{table:spec-errors}.  All of
them were confirmed by TC39 and will be fixed in the next release.  The BigInt
proposal added a new type of primitives BigInt and unified the type with
original Number type to Numeric types.  Thus, it not only added new
algorithms for BigInt but also modified all existing algorithms related to
number values.  The error BigInt-1 is just a simple misuse of the variable
\code{oldValue} in {\bf Evaluation} of {\it UpdateExpression}. BigInt-2 and
BigInt-3 break the backward compatibility because of the incorrectly mixed uses
of {\bf ToInt32} and {\bf ToUint32} algorithms in unsigned right shift operators
and bitwise operators, respectively. The last one BigInt-4 arises because the
Number constructor does not consider BigInt primitives.  On average, ECMAScript
included these four specification errors in \inred{XXX} days.

After fixing all specification errors, we extracted the semantics for the BigInt
proposal again.  The updated semantics passed all 207 applicable tests for the
proposal and \inred{XXXXX} applicable tests for ES10.  Thus, \( \tool \) also
correctly extract IR-based semantics from future proposals.