### 2020-04-30 ase20/conclusion.tex
Till Sec. 7

\section{Conclusion}\label{sec:conclude}
Annual updates of ECMAScript make it difficult to build program
analysis or formal verification of JavaScript due to the required human efforts
in modeling a moving target.  In this paper, we proposed \( \tool \), a tool
that \textit{automatically} extracts the syntax and semantics as a parser and
an AST-IR translator from ECMAScript written in English.
The tool automatically extracts all the syntax and \inred{XX.XX\%} of the
semantics for the most recent four versions of ECMAScript (ES7 to ES10).
We evaluated the correctness of the tool by testing the extracted
semantics from ES10 with Test262, the official conformance suite.
Using \inred{X,XXX} failed tests, we found nine specification errors,
four of which are newly discovered, confirmed by TC39, and planned to
be integrated in ES11.  After fixing the errors, the extracted
semantics passed all \inred{XX,XXX} applicable tests in Test262.
We also showed that \( \tool \) is forward compatible by applying it to nine
proposals to be included in ES11, which let us find four errors in the
BigInt proposal.  We believe that \( \tool \) can dramatically reduce
human efforts in building various JavaScript tools correctly.