### 2021-03-29 paper/ase21/related.tex
[ASE21] Format

\section{Related Work}\label{sec:related}

\begin{itemize}
  \item JISET: JavaScript IR-based Semantics Extraction Toolchain~\cite{jiset}
\end{itemize}