### 2021-04-06 paper/ase21/abstract.tex
[ASE21] Intro

\begin{abstract}
  JavaScript is one of mainstream programming languages not only in client-side
  programming but also in server-side programming and even embedded systems.
  Thus, various JavaScript engines are developed and maintained in various
  fields and they must conform to the syntax and semantics described in
  ECMAScript, the standard specification of JavaScript.  Since an incorrect
  description in ECMAScript can lead to the wrong implementation of JavaScript
  engines, it is important to check the correctness of ECMAScript.  However, all
  the specification updates are currently manually checked by the Ecma Technical
  Committee 39 (TC39) without any automated tools.  Moreover, in late 2014, the
  committee announced yearly release cadence and open development process of
  ECMAScript to quickly adapt to an evolving development environment.  Thus, the
  manual checking becomes more labor-intensive and error-prone.

  To alleviate this problem, we propose $\tool$, a JavaScript Specification Type
  Analyzer using Refinement.  It is the first tool that performs \textit{type
  analysis} on JavaScript specifications and detects specification bugs using
  \textit{bug detector}.  For a given specification, our tool first compiles
  each abstract algorithm written in a structured natural language to the
  corresponding function of $\ires$, which is an untyped intermediate
  representation for ECMAScript.  Then, it performs type analysis for compiled
  functions with specification types defined in ECMAScript.  Based on the result
  of type analysis, $\tool$ detects specification bugs using the bug detector
  consisting of four different checkers.  Moreover, to increase precision of
  type analysis, we also present \textit{condition-based refinement} for type
  analysis, which prunes out infeasible abstract states by using conditions of
  assertions and branches.  We evaluated $\tool$ with all \inred{XXX} different
  versions existed in the official ECMAScript repository for the recent
  \inred{three years from 2018 to 2021}.  Our tool detected \inred{XXX}
  specification bugs including only \inred{XXX} false alarms caused by the
  imprecision of type analysis in \inred{XX} minutes on average.  Among
  \inred{XXX} true alarms, \inred{XXX} bugs are resolved after existing for
  \inred{XXX} days on average and \inred{XX} bugs still exist in the latest
  version of ECMAScript for \inred{XXX} days.  We reported all \inred{XX} newly
  found bugs and they are confirmed by the committee.
\end{abstract}