### 2021-04-06 paper/ase21/related.tex

\section{Related Work}\label{sec:related}

\begin{itemize}
  \item Abstract Interpretation~\cite{ai1977, ai1992}
  \item JISET: JavaScript IR-based Semantics Extraction Toolchain~\cite{jiset}
  \item JavaScript Manual Mechanized Specification:
    \begin{itemize}
      \item The Essence of Javascript~\cite{lambdajs}
      \item A Trusted Mechanised JavaScript Specification~\cite{jscert}
      \item JaVerT: JavaScript Verification Toolchain~\cite{javert}
      \item KJS: A complete formal semantics of JavaScript~\cite{kjs}
      \item JSExplain: A Double Debugger for JavaScript~\cite{jsexplain}
      \item JEST: N+ 1-version Differential Testing of Both JavaScript Engines
        and Specification~\cite{jest}
    \end{itemize}
  \item Automatic Mechanized Specification for others:
    \begin{itemize}
      \item Automatic extraction of x86 formal semantics from its natural
        language description~\cite{x86}
      \item Formal semantics extraction from natural language specifications for
        ARM~\cite{arm}
      \item SpecTest: Specification-Based Compiler Testing~\cite{spectest}
      \item BASESPEC: Comparative Analysis of Baseband Software and Cellular
        Specifications for L3 Protocols~\cite{basespec}
    \end{itemize}
\end{itemize}