### 2021-04-08 paper/ase21/checker.tex
[ASE21] Overview

\section{Bug Detector}\label{sec:checker}

\subsection{Reference Checker}
\begin{itemize}
  \item unknown variables
  \item already defined variables
\end{itemize}

\subsection{Parameter Arity Checker}
\begin{itemize}
  \item remaining normal parameters
    (parameter types: normal / optional / variadic)
  \item remaining arguments
\end{itemize}

\subsection{Assertion Checker}
\begin{itemize}
  \item typeof assertions
  \item instance-of assertions
  \item other assertions
\end{itemize}

\subsection{Operand Type Checker}
\begin{itemize}
  \item implicit abrupt completions
  \item typed operators
\end{itemize}