### 2021-04-13 paper/ase21/checker.tex

\section{Bug Detector}\label{sec:checker}

\begin{table}
  \centering
  \caption{Type-related specification bugs fixed by pull requests for the recent
  three years from 2018 to 2021.}
  \label{table:pr-bugs}
  \vspace*{-1.5em}
  \[
    \begin{array}{c|l|r}
      \multicolumn{1}{c}{\textbf{Category}} &
      \multicolumn{1}{|c}{\textbf{Bug Kind}} &
      \multicolumn{1}{|c}{\textbf{\# Bugs}}\\
      \hline

      \multirow{2}{*}{\text{Reference}}
      & \text{Unknown Variables} & \inred{XXX}\\\cline{2-3}
      & \text{Already Defined Variables} & \inred{XXX}\\\hline

      \multirow{1}{*}{\text{Parameter}}
      & \text{Arity Mismatches} & \inred{XXX}\\\hline

      \multirow{1}{*}{\text{Assertion}}
      & \text{Assertion Failures} & \inred{XXX}\\\hline

      \multirow{2}{*}{\text{Type}}
      & \text{Non-Numeric Operands} & \inred{XXX}\\\cline{2-3}
      & \text{Unchecked Abrupt Completions} & \inred{XXX}\\\hline

      \multicolumn{2}{c|}{\textbf{Total}} & \inred{XXX}\\

    \end{array}
  \]
  \vspace*{-1.5em}
\end{table}

We develop a \textit{bug detector} to statically detect type-related
specification bugs in ECMAScript using an augmented abstract transfer
$\detector$ with additional checkers.  The abstract semantics $\asem{\prog}$ of
a program $\prog = (\getfunc, \getinst, \getnext)$ is the least fixpoint of the
original abstract transfer $\atransfer$.  Thus, $\detector(\asem{\prog})$ also
produces $\asem{\prog}$ and detects type-related specification bugs.  Before
implementing checkers, we manually investigate pull requests for the recent
\inred{three years from 2018 to 2021}.  As described in
Table~\ref{table:pr-bugs}, pull requests include \inred{XXX} type-related
specification bugs classified in six different kinds and four categories.  We
implement four checkers for each category bugs: \textit{reference checker},
\textit{arity checker}, \textit{assertion checker}, and \textit{operand type
checker}.  Now, we explain how each checker detects corresponding bugs in the
augmented abstract transfer $\detector$.

\subsection{Reference Checker}

In ECMAScript abstract algorithms, variables are dynamically introduced in any
contexts.  The reference bugs are caused when trying to access not yet defined
variables or to redefine already defined variables.  Such reference bugs are the
most prevalent type-related specification bugs according to the our manual
investigation of previous pull requests.  Thus, we implement the reference
checker by adding additional checks to abstract semantics of variable lookups
$\aseme{\x}$ and variable declarations $\asemi{\kwlet \; \x = \expr}$ to detect
unknown variables and already defined variables, respectively:
\[
  \begin{array}{r@{~}c@{~}l}
    \aseme{\x}(\aenv) &=& \left\{
      \begin{array}{ll}
        \text{unknown variable} \; $\x$
        & \text{if} \; \asemr{\x}(\aenv) = \{ \tabsent \}\\
        \cdots
        & \text{otherwise}\\
      \end{array}
    \right.\\

    \asemi{\kwlet \; \x = \expr}(\lab, \tys)(\aelem) &=& \left\{
      \begin{array}{ll}
        \text{already defined variable} \; $\x$
        & \text{if} \; \aty = \{ \true \}\\
        \cdots
        & \text{otherwise}\\
      \end{array}
    \right.\\ &&
    \text{where} \; \aty = \aseme{\x \kwexists}(\aelem(\lab, \tys))\\
  \end{array}
\]
If the abstract semantics of a variable lookup for $\x$ is a singleton $\{
\tabsent \}$, it means the variable $\x$ is always an unknown variable.  In this
case, the reference checker reports a unknown variable bug for $\x$.  For
example, revisit the syntax-directed algorithm at the upper-left in
Figure~\ref{fig:example}.  Since the \textbf{GetReferencedName} algorithm is
removed, the variable \code{GetReferencedName} does not exist in abstract
environments and its lookup always return $\{ \tabsent \}$ thus the reference
checker can detect this unknown variable bug.  For already defined variables,
the reference checker examines whether the variable of each variable declaration
is always defined using the abstract semantics of existence checks $\aseme{\x
\kwexists}$.

\subsection{Arity Checker}
Arity $\size{\func}$ of a function $\func = \kwdef \; \f (\p_1, \cdots, \p_n,
\kwsl \cdots, p_m \kwsr). \; \lab$ is defined as an interval $[n, m]$ where $n$
and $m-n$ denote numbers of normal and optional parameters, respectively.  The
arity checker detects mismatches of such arities with the number of arguments by
adding an additional checks in the function call instructions $\asemi{\x = \kwrl
\expr_0 \; \expr_1 \cdots \expr_k \kwrr}$:
\[
  \begin{array}{l}
    \asemi{\x = \kwrl \expr_0 \; \expr_1 \cdots \expr_k \kwrr}(\lab,
    \tys)(\aelem) =\\
    \qquad \left\{
      \begin{array}{ll}
        \text{missing parameters} \; \p_{k+1}, \cdots, \p_{n_\func} &
        \text{if} \; \exists \func \in \aty_0. \; \text{s.t.} \; k < n_\func\\

        \text{reamining arguments} \; \expr_{m_\func+1}, \cdots, \expr_k &
        \text{if} \; \exists \func \in \aty_0. \; \text{s.t.} \; k > m_\func\\

        \cdots &
        \text{otherwise}\\
      \end{array}
    \right.\\
  \end{array}
\]
where $\func = \kwdef \; \f (\p_1, \cdots, \p_{n_\func}, \kwsl \cdots,
p_{m_\func} \kwsr). \; \lab \wedge \aty_0 = \aseme{\expr_0}(\aelem(\lab,
\tys))$.  For each function $\func$ in the abstract semantics of the function
expression $\expr_0$, the arity checker compares the number of arguments with
its arity $\size{\func}$ and reports a corresponding arity bug.  For example,
consider the lower-left syntax directed algorithm in Figure~\ref{fig:example}.
The algorithm invocation in line 2 is compiled to a function call instruction:
\[
  \x = \kwrl \code{formal}.\code{IteratorBindingInitialization} \; \code{formal}
  \kwrr
\]
with a temporal variable $\x$ thus the number of arguments is 1.  However, arity
of the function stored in $\code{formal}.\code{IteratorBindingInitialization}$
is $[3, 3]$ thus the arity checker reports missing parameter bugs for two
additional parameters $\code{iteratorRecord}$ and $\code{environment}$.

\subsection{Assertion Checker}
\todo

\subsection{Operand Type Checker}
\todo