### 2021-04-14 paper/ase21/related.tex
[ASE21] temporal dump

\section{Related Work}\label{sec:related}

\paragraph{ECMAScript-based Applications}
\begin{itemize}
  \item Engines~\cite{v8, graaljs, qjs, moddable}
  \item Static Analyzer~\cite{safe, tajs, jsai, wala}
  \item Debuggers: JSExplain: A Double Debugger for JavaScript~\cite{jsexplain}
  \item Verification Tool: JaVerT: JavaScript Verification
    Toolchain~\cite{javert}
\end{itemize}

\paragraph{Mechanized Specifications}
\begin{itemize}
  % manual
  \item The Essence of Javascript~\cite{lambdajs}
  \item A Trusted Mechanised JavaScript Specification~\cite{jscert}
  \item KJS: A complete formal semantics of JavaScript~\cite{kjs}

  % automatic
  \item BASESPEC: Comparative Analysis of Baseband Software and Cellular
    Specifications for L3 Protocols~\cite{basespec}
  \item Automatic extraction of x86 formal semantics from its natural
    language description~\cite{x86}
  \item Formal semantics extraction from natural language specifications for
    ARM~\cite{arm}
  \item SpecTest: Specification-Based Compiler Testing~\cite{spectest}
  \item Automatic Model Generation from Documentation for Java API Functions
  \item COMFORT: JS Fuzzer using ECMAScript
  \item JISET: JavaScript IR-based Semantics Extraction Toolchain~\cite{jiset}
\end{itemize}

\paragraph{Correctness of ECMAScript}
\begin{itemize}
  \item JISET: JavaScript IR-based Semantics Extraction Toolchain~\cite{jiset}
  \item JEST: N+1-version Differential Testing of Both JavaScript Engines
    and Specification~\cite{jest}
\end{itemize}

% \begin{itemize}
%   \item Abstract Interpretation~\cite{ai1977, ai1992}
%   \item JISET: JavaScript IR-based Semantics Extraction Toolchain~\cite{jiset}
%   \item JavaScript Manual Mechanized Specification:
%     \begin{itemize}
%       \item The Essence of Javascript~\cite{lambdajs}
%       \item A Trusted Mechanised JavaScript Specification~\cite{jscert}
%       \item JaVerT: JavaScript Verification Toolchain~\cite{javert}
%       \item KJS: A complete formal semantics of JavaScript~\cite{kjs}
%       \item JEST: N+ 1-version Differential Testing of Both JavaScript Engines
%         and Specification~\cite{jest}
%     \end{itemize}
%   \item Automatic Mechanized Specification for others:
%     \begin{itemize}
%       \item Automatic extraction of x86 formal semantics from its natural
%         language description~\cite{x86}
%       \item Formal semantics extraction from natural language specifications for
%         ARM~\cite{arm}
%       \item SpecTest: Specification-Based Compiler Testing~\cite{spectest}
%       \item BASESPEC: Comparative Analysis of Baseband Software and Cellular
%         Specifications for L3 Protocols~\cite{basespec}
%     \end{itemize}
% \end{itemize}