### 2021-04-18 paper/ase21/related.tex

\section{Related Work}\label{sec:related}

JavaScript specification type analysis is related to two research fields:
JavaScript tools, mechanized specification extraction, specification-based
testing.

\paragraph{JavaScript Tools}
ECMAScript is the standard language specification for JavaScript maintained by
the Ecma Technical Committee 39 (TC39).  In late 2014, the committee announced
yearly release cadence and open development process of ECMAScript to quickly
adapt to an evolving development environment.  Existing JavaScript engines, such
as Google V8~\cite{v8}, GraalJS~\cite{graaljs}, QuickJS~\cite{qjs}, and Moddable
XS~\cite{moddable}, must conform to the syntax and semantics described in
annually updated ECMAScript.  Beyond JavaScript engines, there are diverse
research directions to utilize JavaScript specifications.  Several researchers
developed static analyzers, such as TAJS~\cite{tajs}, SAFE~\cite{safe},
JSAI~\cite{jsai}, and WALA~\cite{wala}, based on abstract interpretation
framework~\cite{ai1977, ai1992} with various analysis techniques.  They defined
abstract semantics of JavaScript semantics described in ECMAScript to statically
analyze given JavaScript programs in a finite time.  \citet{jsexplain} presented
JSExplain, a debugger for JavaScript, by implementing a reference interpreter in
OCaml directly conforms to algorithm steps in ECMAScript.  For a given
JavaScript program, the debugger produces execution traces interactively
invesitgated in a browser, with an interface that displays not only the code and
the state of the interpreter.  \citet{javert} introduced JaVerT, a JavaScript
verification toolchain, based on separation logic with their own intermediate
goto language JSIL.  Moreover, JaVerT 2.0~\cite{javert2} extends it to support
compositional symbolic execution for JavaScript based on bi-abduction.  However,
all of them manually handles ECMAScript with their own intermediate
representations thus many of them still targets ES5.1, the most famous old
version of ECMAScript released in 2011, not the latest one.

\paragraph{Mechanized Specification Extraction}
Several researchers extracted mechanized specifications from specifications
written in a natural language to mechanically handle materials in
specifications.  For system architectures, researchers utilized complex Natural
Language Processing (NLP) and Machine Learning (ML) to extract formal semantics
of small-sized low-level assembly languages for x86~\cite{x86} and
ARM~\cite{arm}.  For Java API functions, \citet{javadoc} presented a technique
to automatically extract models from their documentation using NLP techniques.
Instead of specific API functions, \citet{jiset} presented $\jiset$, a tool that
extracts mechanized language specification from ECMAScript.  While previous
JavaScript formal semantics~\cite{lambdajs, jscert, kjs} were manually defined,
$\jiset$ automatically extracts formal semantics directly from ECMAScript.  We
utilized $\jiset$ to perform type analysis for diverse versions of ECMAScript in
$\tool$.

\paragraph{Specification-based Testing}
Several researchers utilized specifications to test their implementations.  For
network protocols, \citet{basespec} proposed a novel approach named
\textsc{BaseSpec}, which extracts message structures from tables in cellular
specifications for L3 protocols to perform comparative analysis with baseband
software.  \citet{spectest} presented SpecTest that utilized an executable
language semantics to perform fuzzing for Java and Solidity compilers.  For
JavaScript, \citet{comfort} presented \textsc{Comfort}, a compiler fuzzing
framework to detect JavaScript engine bugs using ECMAScript with deep
learning-based language models.  \citet{jest} extended $\jiset$ to $\jest$ that
performs $N$+1-version differential testing with a reference interpreter
extracted from ECMAScript and $N$ different JavaScript engines.  $\jest$ detects
not only engine bugs but also specification bugs in ECMAScript based on the
cross-referencing oracle.  However, it requires multiple JavaScript engines and
takes dozens of hours to cover semantics for each version of ECMAScript.
Instead, $\tool$ can detect specification bugs without JavaScript engines less
than 5 minutes.  Moreover, $\tool$ utilizes abstract semantics instead of
concrete semantics used in $\jest$ thus it can quickly cover more semantics.