### 2021-04-19 paper/ase21/overview.tex
[ASE21] Updated overview

\section{Overview}\label{sec:overview}

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{img/overall}
  \caption{Overall structure of $\tool$: a type analyzer and a bug detector for
  mechanized specifications extracted from ECMAScript by $\jiset$.}
  \label{fig:overall}
  \vspace*{-1.5em}
\end{figure}

\begin{figure*}[t]
  \centering
  \includegraphics[width=\textwidth]{img/example}
  \vspace*{-1.5em}
  \caption{An example JavaScript program with related previous specification
  bugs and their bugfixes.}
  \label{fig:example}
  \vspace*{-1.5em}
\end{figure*}
% XXX
% \begin{lstlisting}[style=FigureJS]
% let f;    f ??= (x) => Math.round(x);    f(true);
% \end{lstlisting}

In this section, we demonstrate overall structure of $\tool$ depicted in
Figure~\ref{fig:overall}.  Our tool consists of three different steps:
1) specification extraction, 2) type analysis, and 3) bug detection.

\subsection{Specification Extraction}\label{sec:overview-spec-extract}

ECMAScript describes JavaScript syntax and semantics written in a structured
natural language.  We utilize $\jiset$ to extract mechanized specification from
a given ECMAScript.

\paragraph{Syntax and Semantics} ECMAScript uses a variant of extended BNF
notations to describe JavaScript syntax.  $\jiset$ synthesizes AST structures
and JavaScript parsers from its lexical and syntactic productions.  Our tool
utilizes only AST structures to define their types and to access corresponding
syntax-directed algorithms.  For JavaScript semantics, ECMAScript contains
abstract algorithms written in a structured natural language.  $\jiset$ compiles
them to functions of $\ires$, an untyped intermediate representation for
ECMAScript.  We modify $\ires$ to make it more suitable for type analysis, and
formally define it in Section~\ref{sec:ires}.

\paragraph{Types} In ECMAScript, types denote not only JavaScript types but also
specification types for ASTs, records, and list-like structures.  In abstract
algorithms, ASTs are also values thus they can be stored in the variables or
passed as arguments.  For such ASTs, we use their production names as their
types and automatically link corresponding syntax-directed algorithms to them as
their fields.  ECMAScript also supports various record types and their fields
and possible values are defined in the corresponding tables.  For example,
``Table 9: Completion Record Fields'' in the latest ECMAScript describes fields
of the completion records.  The latest version contains tens of tables for
record types but they are backward compatible in general.  Thus, we manually
model fields of record types based on tables in the latest version and use use
them in type analysis.  For list-like structures, we define the empty list type
$\clist{}$ and parametric list types $\clist{\tau}$.

\subsection{Type Analysis}\label{sec:overview-type-analysis}

The type analyzer performs type analysis with flow-sensitivity and
type-sensitivity for arguments to precisely detect specification bugs.  Each
function is divided into multiple flow- and type-sensitive views, and an
abstract state stores mapping from views to corresponding abstract environments.
To separately handle views, we use a worklist algorithm.  The type analyzer
consists of two sub-modules: an \textit{analysis initializer} and an
\textit{abstract transfer}.

\paragraph{Analysis Initializer} It defines the initial abstract state and the
initial set of views in worklist.  In ECMAScript, there are three different
kinds of abstract algorithms \textit{normal}, \textit{syntax-directed} and
\textit{built-in}.  Since all the parameter types are known for built-in
algorithms and syntax-directed algorithms having no additional parameters, we
utilize them as entry points of type analysis.  For each entry point, the
initializer defines its abstract environment with parameter types and pushes the
flow- and type-sensitive views of entry points to the worklist.

\paragraph{Abstract Transfer} For each iteration, the abstract transfer gets a
specific view from the worklist and udpates the abstract environment of next
views based on an abstract semantics.  It pushes next views to the worklist if
it changes their abstract environments, and the iteration will be finished when
the worklist becomes empty.  Moreover, to increase the analysis precision, we
perform condition-based refinement for abstract environment when the current
control point is a branch or an assertion.  The detail of refinement will be
described in Section~\ref{sec:refine}.

\subsection{Bug Detection}\label{sec:overview-bug-detect}

To detect specification bugs based on the result of type analysis, we develop
four different checkers in the bug detector.  We explain targets of checkers
using an example JavaScript program with related previous specification bugs and
their bugfixes descirbed in Figure~\ref{fig:example}.

\paragraph{Reference Checker} The example JavaScript program first defines a
variable \jscode{f} without initialization thus it stores \jscode{undefined}.
Then, it assigns an arrow function to \jscode{f} using the operator
\jscode{??=}.  While the corresponding \textbf{Evaluation} algorithm originally
used \textbf{GetReferencedName} algorithm to get reference name in line 4.a, a
contributor removed the \textbf{GetReferencedName} algorithm and replaced its
all invocations to accesses of the field [[ReferencedName]] on October 28, 2020.
However, there were several missing cases including the semantics of
\jscode{??=} until another contributor fixed this bug on November 3, 2020.
Thus, an unknown variable bug for \textbf{GetReferencedName} existed for 7 days.
The reference checker detects reference bugs including such unknown variable
bugs.

\paragraph{Arity Checker} After the program assigns an arrow function to
\jscode{f}, it calls the function with an argument \jscode{true}.  During the
initialization of a function call, the \textbf{IteratorBindingInitialization}
algorithm is evaluated with additional parameters \textit{iteratorRecord} and
\textit{environment} to assign arguments to proper parameters.  However, a
contributor missed to pass additional arguments for them in line 2 in the
\textbf{IteratorBindingInitialization} of \textit{ArrowParameters} on September
6, 2018.  It caused an arity mismatch bug and existed for 533 days until another
contributor fixed this bug on February 20, 2020.  The arity checker checks such
arity mismatches.

\paragraph{Assertion Checker} During the initialization of the function call,
another specification bug existed in \textbf{IteratorBindingInitialization} of
\textit{FormalParameter}.  The additional \textit{environment} parameter might
contains \jscode{undefined} or an environment record exactly same with
\textit{originalEnv} as described in the assertion at line 5.  However, the
assertion did not consider \jscode{undefined} introduced by the initial commit
of the open development process on September 22, 2015.  Thus, the assertion
failure bug existed for 1,297 days until April 10, 2019.  The assertion checker
checks whether such assertions are failed or not.

\paragraph{Operand Checker} After the function call initialization, the
parameter \jscode{x} points to \jscode{true} thus \jscode{Math.round} is invoked
with the argument \jscode{true}.  The \jscode{Math.round} built-in library first
converts the given parameter \textit{x} to the corresponding number value
\textit{n} using \textbf{ToNumber} algorithm, and then should perform remaining
steps using the number value \textit{n}.  However, a contributor made a mistake
of using \textit{x} instead of \textit{n} in line 3 and 4 on September 11, 2020.
It makes the specification tries to compare a boolean value \jscode{true} with a
numeric value 0.5 or 0 in the example JavaScript program.  Fortunately, another
contributor fixed this bug after two days.  The operand checker detects such
bugs by checking whether the operand types are compatible with the typed
operators.

In the remainder of this paper, we explain the details of how to perform type
analysis for $\ires$ functions and how to increase analysis precision using
condition-based refinement (Section~\ref{sec:analyzer}) and how to detect
type-related specification bugs using four different checkers
(Section~\ref{sec:checker}).  After we evaluate $\tool$
(Section~\ref{sec:eval}), we discuss related work (Section~\ref{sec:related}),
and conclude (Section~\ref{sec:conclusion}).