### 2021-04-20 paper/ase21/conclusion.tex
[ASE21] Till conclusion

\section{Conclusion}\label{sec:conclusion}
Since an incorrect description in ECMAScript can lead to the wrong JavaScript
engine implementations, checking the correctness of ECMAScript is essential.
Since ECMAScript is annually released and developed in open process, it becomes
more labor-intensive and error-prone.  To alleviate the problem, we propose
$\tool$, a JavaScript Specification Type Analyzer using Refinement.  It is the
first tool that performs \textit{type analysis} on JavaScript specifications and
detects specification bugs using a \textit{bug detector}.  Moreover, to increase
precision of type analysis, we present \textit{condition-based refinement} for
type analysis, which prunes out infeasible abstract states using conditions of
assertions and branches.  We evaluated $\tool$ with all 864 versions in the
official ECMAScript repository for the recent three years from 2018 to 2021.
Our tool took \inred{136.0} seconds on average to perform type analysis for each
version, and detected \inred{177} type-related specification bugs with
\inred{52.0}\% precision; \inred{92} out of \inred{177} bugs are true bugs.
Among them, \inred{14} bugs are newly detected by $\tool$, and the committee
confirmed all of them.