### 2021-04-21 paper/ase21/related.tex
[ASE21] Removed natbib package

\section{Related Work}\label{sec:related}

JavaScript specification type analysis is related to two research fields:
JavaScript tools, mechanized specification extraction, specification-based
testing.

\paragraph{JavaScript Tools}
ECMAScript is the standard language specification for JavaScript maintained by
the Ecma Technical Committee 39 (TC39).  In late 2014, the committee announced
yearly release cadence and open development process of ECMAScript to quickly
adapt to an evolving development environment.  Existing JavaScript engines, such
as Google V8~\cite{v8}, GraalJS~\cite{graaljs}, QuickJS~\cite{qjs}, and Moddable
XS~\cite{moddable}, must conform to the syntax and semantics described in
annually updated ECMAScript.  Beyond JavaScript engines, there are diverse
research directions to utilize JavaScript specifications.  Several researchers
developed static analyzers, such as TAJS~\cite{tajs}, SAFE~\cite{safe},
JSAI~\cite{jsai}, and WALA~\cite{wala}, based on abstract interpretation
framework~\cite{ai1977, ai1992} with various analysis techniques.  They defined
abstract semantics of JavaScript semantics described in ECMAScript to statically
analyze given JavaScript programs in a finite time.  Chargu{\'e}raud et
al.~\cite{jsexplain} presented JSExplain, a debugger for JavaScript, by
implementing a reference interpreter in OCaml directly conforms to algorithm
steps in ECMAScript.  For a given JavaScript program, the debugger produces
execution traces interactively invesitgated in a browser, with an interface that
displays not only the code and the state of the interpreter.  Fragoso Santos et
al.~\cite{javert} introduced JaVerT, a JavaScript verification toolchain, based
on separation logic with their own intermediate goto language JSIL.  Moreover,
JaVerT 2.0~\cite{javert2} extends it to support compositional symbolic execution
for JavaScript based on bi-abduction.  However, all of them manually handles
ECMAScript with their own intermediate representations thus many of them still
targets ES5.1, the most famous old version of ECMAScript released in 2011, not
the latest one.

\paragraph{Mechanized Specification Extraction}
Several researchers extracted mechanized specifications from specifications
written in a natural language to mechanically handle materials in
specifications.  For system architectures, researchers utilized complex Natural
Language Processing (NLP) and Machine Learning (ML) to extract formal semantics
of small-sized low-level assembly languages for x86~\cite{x86} and
ARM~\cite{arm}.  For Java API functions, Zhai et al.~\cite{javadoc} presented a
technique to automatically extract models from their documentation using NLP
techniques.  Instead of specific API functions, Park et al.~\cite{jiset}
presented $\jiset$, a tool that extracts mechanized language specification from
ECMAScript.  While previous JavaScript formal semantics~\cite{lambdajs, jscert,
kjs} were manually defined, $\jiset$ automatically extracts formal semantics
directly from ECMAScript.  We utilized $\jiset$ to perform type analysis for
diverse versions of ECMAScript in $\tool$.

\paragraph{Specification-based Testing}
Several researchers utilized specifications to test their implementations.  For
network protocols, Kim et al.~\cite{basespec} proposed a novel approach named
\textsc{BaseSpec}, which extracts message structures from tables in cellular
specifications for L3 protocols to perform comparative analysis with baseband
software.  Schumi and Sun~\cite{spectest} presented SpecTest that utilized an
executable language semantics to perform fuzzing for Java and Solidity
compilers.  For JavaScript, Ye et al.~\cite{comfort} presented \textsc{Comfort}, a
compiler fuzzing framework to detect JavaScript engine bugs using ECMAScript
with deep learning-based language models.  Park et al.~\cite{jest} extended
$\jiset$ to $\jest$ that performs $N$+1-version differential testing with a
reference interpreter extracted from ECMAScript and $N$ different JavaScript
engines.  $\jest$ detects not only engine bugs but also specification bugs in
ECMAScript based on the cross-referencing oracle.  However, it requires multiple
JavaScript engines and takes dozens of hours to cover semantics for each version
of ECMAScript.  Instead, $\tool$ can detect specification bugs without
JavaScript engines less than 5 minutes.  Moreover, $\tool$ utilizes abstract
semantics instead of concrete semantics used in $\jest$ thus it can quickly
cover more semantics.