### 2021-04-21 paper/ase21/report/abstract.tex
[ASE21] Init report

\begin{abstract}
  In this report, we formally define a modified $\ires$ with its syntax and
  semantics.  Moreover, we define type analysis for the modified $\ires$.
\end{abstract}