### 2021-04-21 paper/ase21/report/syntax.tex

\section{Syntax}\label{sec:syntax}
We first define syntax of the modified $\ires$ as follows:
\[
  \vspace*{-1em}
  \begin{array}{l@{~}r@{~}c@{~}r@{~}r@{~}l}
    \text{Functions}
    &\funcset&\ni&\func&::=&\kwdef \; \x (\x^*, \kwsl \x^* \kwsr). \; \lab\\

    \text{Instructions}
    &\instset&\ni&\inst&::=&
    \kwlet \; \x = \expr \mid
    \x = \kwrl \expr \; \expr^* \kwrr \mid
    \kwassert \; \expr \\

    &&&&\mid&
    \kwif \; \expr \; \lab \; \lab \mid
    \kwreturn \; \expr \mid
    \refer = \expr \\

    \text{References}
    &&&\refer&::=&
    \x \mid
    \refer \kwsl \expr \kwsr \\

    \text{Expressions}
    &&&\expr&::=&
    \tname \; \kwcl [\x: \expr]^* \kwcr \mid
    \clist{\expr^*} \mid
    \expr: \ty \mid
    \refer \kwexists \\

    &&&&\mid&
    \expr \bop \expr \mid
    \uop \; \expr \mid
    \refer \mid
    \const \mid
    \prim \\

    \text{Primitives}
    &\primset&\ni&\prim&::=&
    \undefval \mid \nullval \mid \bool \mid
    \num \mid \bigint \mid \str \mid \symb \\

    \text{Types} &\tyset&\ni&\ty&::=&\tname \mid \clist{} \mid \clist{\ty} \mid
    \tjs \mid \tprim\\

    &&&&\mid&
    \undefval \mid \nullval \mid \tbool \mid \tnumeric\\

    &&&&\mid&
    \tnum \mid \tbigint \mid \tstr \mid \tsymb \\
    \vspace*{1em}
  \end{array}
\]