### 2021-04-21 paper/ase21/conclusion.tex
Revised Sec. 6 & 7

\section{Conclusion}\label{sec:conclusion}
Checking the correctness of ECMAScript is essential because an incorrect description
in ECMAScript can lead to wrong implementations of JavaScript engines. However,
since ECMAScript is annually released and developed in an open process, checking
its correctness becomes more labor-intensive and error-prone.  To alleviate the problem,
we propose $\tool$, a JavaScript Specification Type Analyzer using Refinement.
It is the first tool that performs \textit{type analysis} on JavaScript specifications and
detects specification bugs using a \textit{bug detector}.  We also present
\textit{condition-based refinement} for type analysis, which prunes out infeasible abstract states
using conditions of assertions and branches to improve the analysis precision.
We evaluated $\tool$ with all 864 versions in the official ECMAScript repository
for the last three years from 2018 to 2021.  It took \inred{137.3} seconds on average
to perform type analysis for each version, and detected \inred{161} type-related specification bugs with
\inred{56.5}\% precision; \inred{91} out of \inred{161} reported bugs are true bugs.
Among them, \inred{14} bugs are newly detected by $\tool$, and the
committee confirmed them all.