### 2021-04-21 paper/ase21/related.tex

\section{Related Work}\label{sec:related}

Type analysis of JavaScript specifications has three related topics:
JavaScript tools, mechanized specification extraction, and specification-based testing.

\paragraph{JavaScript Tools}
ECMAScript is the standard language specification for JavaScript maintained by
the Ecma Technical Committee 39 (TC39).  In late 2014, the committee announced
its plan to release ECMAScript annually and adopt the open development process
to quickly adapt to evolving development environments.  Various JavaScript engines such
as Google V8~\cite{v8}, GraalJS~\cite{graaljs}, QuickJS~\cite{qjs}, and Moddable
XS~\cite{moddable} should conform to the syntax and semantics described in
annually updated ECMAScript.  Beyond JavaScript engines, diverse
research projects use JavaScript specifications.  The main research direction
has been static analyzers such as JSAI~\cite{jsai}, SAFE~\cite{safe},
TAJS~\cite{tajs}, and WALA~\cite{wala} based on the abstract interpretation
framework~\cite{ai1977, ai1992} with their own analysis techniques.
They defined abstract semantics of the JavaScript semantics described in ECMAScript to statically
analyze JavaScript programs in a finite time.
Chargu{\'e}raud et al.~\cite{jsexplain} presented JSExplain, a debugger for JavaScript, by
implementing a reference interpreter in OCaml following the algorithm steps in ECMAScript closely.
For a given JavaScript program, the debugger interactively produces execution traces
investigated in a browser, with an interface that displays the
JavaScript code and the interpreter's state.
Fragoso Santos et al.~\cite{javert} introduced JaVerT, a JavaScript verification toolchain, based
on the separation logic with an intermediate goto language JSIL.
JaVerT 2.0~\cite{javert2} extends it to support compositional symbolic execution
for JavaScript based on bi-abduction.  However, because all of them manually handle
ECMAScript with their own intermediate representations, most of them still 
target ES5.1 released in 2011 instead of the latest one.

\paragraph{Mechanized Specification Extraction}
Researchers in various application domains have extracted mechanized specifications from
specifications written in natural languages to handle the contents in the specifications automatically.
For system architectures, researchers utilized Natural
Language Processing (NLP) and Machine Learning (ML) techniques to extract formal semantics
of small-sized low-level assembly languages for x86~\cite{x86} and ARM~\cite{arm}.
For Java API functions, Zhai et al.~\cite{javadoc} presented a
technique to automatically extract models from their documentation using NLP techniques.
For the JavaScript programming language, Park et al.~\cite{jiset}
presented $\jiset$, a tool that extracts a mechanized specification from ECMAScript.
While all the previous JavaScript formal semantics~\cite{lambdajs, jscert,
kjs} were manually defined, $\jiset$ automatically extracts formal semantics directly from ECMAScript.
We utilized $\jiset$ to analyze 864 ECMAScript versions via $\tool$.

\paragraph{Specification-based Testing}
Recently, researchers have utilized specifications to test their implementations.
For network protocols, Kim et al.~\cite{basespec} proposed a novel approach named
\textsc{BaseSpec}, which extracts message structures from tables in cellular
specifications for L3 protocols to perform comparative analysis of baseband software.
Schumi and Sun~\cite{spectest} presented SpecTest, which utilized an
executable language semantics to perform fuzzing for Java and Solidity compilers.
For JavaScript, Ye et al.~\cite{comfort} presented \textsc{Comfort}, a
compiler fuzzing framework to detect JavaScript engine bugs using ECMAScript
with deep learning-based language models.
Park et al.~\cite{jest} extended $\jiset$ to $\jest$, which performs $N$+1-version differential testing
with $N$ different JavaScript engines and a reference interpreter extracted from ECMAScript.
$\jest$ detects not only engine bugs but also specification bugs in
ECMAScript using the cross-referencing oracle.  However, it requires multiple JavaScript engines and
takes dozens of hours to test a version of ECMAScript.
Instead, $\tool$ can detect specification bugs without JavaScript engines in less than five minutes.
Because $\tool$ uses abstract semantics while $\jest$ uses concrete semantics,
$\tool$ can quickly analyze more semantics than $\jest$.