### 2021-04-22 paper/ase21/eval.tex

\section{Evaluation}\label{sec:eval}

\begin{table*}
  \centering
  \caption{The analysis precision of $\tool$ without refinement
  (\stextsf{no-refine}), with refinement (\stextsf{refine}), and their
  difference ($\Delta$)}
  \label{table:precision}
  \resizebox{\textwidth}{!}{%
    \begin{tabular}{c|l?r@{~}c@{~}r@{~}l@{}r|r@{~}c@{~}r@{~}l@{}r?r@{~}c@{~}r@{~}l@{}r|r@{~}c@{~}r@{~}l@{}r?r@{~}c@{~}r@{~}l@{}r|r@{~}c@{~}r@{~}l@{}r}
      \multirow{2}{*}{\textbf{Checker}} &
      \multirow{2}{*}{\textbf{Bug Kind}} &
      \multicolumn{30}{c}{
        \textbf{Precision = (\# True Bugs) / (\# Detected Bugs)}
      }\\\cline{3-32} &&
      \multicolumn{10}{c?}{\stextsf{\footnotesize no-refine}} &
      \multicolumn{10}{c?}{\stextsf{\footnotesize refine}} &
      \multicolumn{10}{c}{$\Delta$}\\

      \specialrule{.1em}{.0em}{.0em}

      \myrowc
      {\mr{2}{\stextsf{\scriptsize Reference}}}
      {\me{2}{62}{106}{58.5}} {\me{2}{63}{78}{80.8}} {\me{2}{+1}{-28}{+22.3}}
      {\mr{1}{\stextsf{\scriptsize UnknownVar}}}
      {\me{1}{17}{60}{28.3}} {\me{1}{17}{31}{54.8}} {\me{1}{}{-29}{+26.5}}

      \myrowh
      {\mr{1}{\stextsf{\scriptsize DuplicatedVar}}}
      {\me{1}{45}{46}{97.8}} {\me{1}{46}{47}{97.9}} {\me{1}{+1}{+1}{+0.1}}

      \hline

      \smyrowc
      {\mr{1}{\stextsf{\scriptsize Arity}}}
      {\mr{1}{\stextsf{\scriptsize MissingParam}}}
      {\me{1}{4}{4}{100.0}} {\me{1}{4}{4}{100.0}} {\me{1}{}{}{}}

      \hline

      \smyrowc
      {\mr{1}{\stextsf{\scriptsize Assertion}}}
      {\mr{1}{\stextsf{\scriptsize Assertion}}}
      {\me{1}{4}{56}{7.1}} {\me{1}{4}{31}{12.9}} {\me{1}{}{-25}{+5.8}}

      \hline

      \myrowc
      {\mr{2}{\stextsf{\scriptsize Operand}}}
      {\me{2}{22}{113}{19.5}} {\me{2}{22}{44}{50.0}} {\me{2}{}{-69}{+30.5}}
      {\mr{1}{\stextsf{\scriptsize NoNumber}}}
      {\me{1}{2}{65}{3.1}} {\me{1}{2}{6}{33.3}} {\me{1}{}{-59}{+30.3}}

      \myrowh
      {\mr{1}{\stextsf{\scriptsize Abrupt}}}
      {\me{1}{20}{48}{41.7}} {\me{1}{20}{38}{52.6}} {\me{1}{}{-10}{+11.0}}

      \specialrule{.1em}{.0em}{.0em}

      \mysrowh
      {\textbf{Total}}
      {92 / 279 (33.0\%)}{93 / 157 (59.2\%)}{+1 / -122 (+26.3\%)}

    \end{tabular}
  }
  \vspace*{-1.5em}
\end{table*}

We implemented $\tool$ as an open-source tool\footnote{The link is anonymized
because of a double-blind review process.} in Scala by extending $\jiset$, a
JavaScript IR-based semantics extraction toolchain~\cite{jiset}, with a worklist-based
fixpoint algorithm for type analysis. Thus, $\tool$ reports type-related specification bugs
detected in fully compiled abstract algorithms by $\jiset$.  For built-in libraries,
$\tool$ analyzes the abstract algorithms of the essential built-in objects: \jscode{Array},
\jscode{Object}, \jscode{Function}, \jscode{Math}, \jscode{Proxy}, and objects
for JavaScript primitive types.

We evaluate $\tool$ using the following research questions:
\begin{itemize}
  \item RQ1. \textbf{(Performance)} How long does $\tool$ take to perform type
    analysis for JavaScript specifications?
  \item RQ2. \textbf{(Precision)} How many type-related specification bugs
    detected by $\tool$ are true bugs?
  \item RQ3. \textbf{(Effectiveness of Refinement)} Does the condition-based refinement
    improve the analysis precision with modest performance overhead?
  \item RQ4: \textbf{(Detection of New Bugs)} Does $\tool$ detect new
    specification bugs in the latest version of ECMAScript?
\end{itemize}
Because the draft of the next ECMAScript (ES12, 2021) is fixed on March 9,
2021, we analyzed all 864 versions in the official
ECMAScript repository\footnote{https://github.com/tc39/ecma262} for the last
three years from January 1, 2018 to March 9, 2021.  We performed our experiments
on five Ubuntu machines equipped with 4.2GHz Quad-Core Intel Core i7 and 32GB of
RAM.

\subsection{Performance}\label{sec:performance}

Figure~\ref{fig:stat} shows the statistics of the type analysis using $\tool$ for
864 versions of ECMAScript: (a) the number of analyzed functions, (b) the number
of flow- and type-sensitive views, (c) the number of worklist iterations, and
(d) the analysis time.  For each version, $\tool$ analyzed
1,696.6 functions on average.  Since ECMAScript has gradually evolved,
it analyzed 1,491 functions for the first version in 2018
but analyzed 1,864 functions in the latest.
$\tool$ analyzes functions with flow- and type-sensitive views.
On average, each version has 92.0K views and each function has
54.1~views.

We measured the performance of $\tool$ with the worklist iteration number
and the analysis time.  For each version of ECMAScript,
$\tool$ took 137.3 seconds with 301.6K worklist iterations
on average.  The average analysis time is 8.0 seconds
for specification extraction (\stextsf{extract}),
128.5 seconds for type analysis (\stextsf{analyze}), and
0.8 seconds for bug detection (\stextsf{detect}).
The performance overhead is modest enough for $\tool$ to be integrated
in the open development process of ECMAScript.

\subsection{Precision}\label{sec:precision}

% \begin{table}
%   \centering
%   % \includegraphics[width=\columnwidth]{img/author}
%   \caption{The creators and resolvers of true bugs.}
%   \label{fig:author}
%   \resizebox{\columnwidth}{!}{$
%     \inred{
%     \begin{array}{c?r|r|r|r|r|r?c?c}
%       \textbf{Rank} >
%       & 0 & 10 & 20 & 30 & 40 & 50
%       & \multicolumn{1}{c?}{\multirow{2}{*}{\textbf{Total}}}
%       & \multicolumn{1}{c}{\multirow{2}{*}{\textbf{Unknown}}}\\\cline{1-7}
%       \textbf{Rank} \leq
%       & 10 & 20 & 30 & 40 & 50 & 60
%       & & \\
% 
%       \specialrule{.1em}{.0em}{.0em}
% 
%       \textbf{\# Created}
%       & 23 & 4 & 1 & 14 & 2 & 1
%       & 45 & 46\\\hline
%       \textbf{\# Resolved}
%       & 57 & 6 & 12 & 2 & 0 & 0
%       & 77 & 14\\
% 
%     \end{array}
%     }
%   $}
%   \vspace*{-1em}
% \end{table}

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.24\textwidth}
    \includegraphics[width=\textwidth]{img/life-sort}
    \caption{Life spans sorted by creation}
  \end{subfigure}
  \begin{subfigure}[b]{0.24\textwidth}
    \includegraphics[width=\textwidth]{img/life-hist}
    \caption{The histogram of life spans}
  \end{subfigure}
  \caption{Life spans of true bugs}
  \vspace*{-1.5em}
  \label{fig:life}
\end{figure}

We measured the analysis precision with the ratio of true bugs in the reported bugs by $\tool$.
As summarized in the \stextsf{refine} column of Table~\ref{table:precision}, the analysis precision
is 59.2\%; 93 out of 157 detected bugs are true bugs.
The reference checker detected the most bugs with 80.8\% precision;
17 unknown variables (\stextsf{UnknownVar}) and
46 duplicated variable declarations (\stextsf{DuplicatedVar}) are true bugs.
We found four missing parameters (\stextsf{MissingParam}) with 100.0\% precision and
four assertion failures (\stextsf{Assertion}) with 12.9\% precision.
Finally, the operand checker detected
two non-numeric operand bugs (\stextsf{NoNumber}) with 33.3\% precision and
20 unchecked abrupt completion bugs (\stextsf{Abrupt}) with 52.6\% precision.

To understand the impact of the detected true bugs, we extended $\tool$ to
automatically extract when they are created and resolved in the ECMAScript
official repository.  A bug is \textit{created} when it exists
in a specific version but does not exist in its previous version, and
a bug is \textit{resolved} vice versa. The \textit{life span} of a bug denotes
the number of days between the created date and the resolved date.
Figure~\ref{fig:life} illustrates the life spans of true bugs;
Figure~\ref{fig:life}(a) depicts the life spans sorted by creation and
Figure~\ref{fig:life}(b) depicts the histogram of the life spans in a logarithmic scale.
Among 93 true bugs, 48 bugs are \textit{inherited},
which means that they are created before 2018,
and 14 bugs still exist in the latest ECMAScript, which are newly detected by $\tool$.
We discuss the details of 14 newly found bugs in Section~\ref{sec:new-bug}.
Even though we assume that 48 inherited bugs are created on January 1, 2018,
the average life span is 414.5 and the maximum life span is 1,164.
All the bugs with the maximum life span are inherited ones and they are all newly detected.

We manually investigated 64 false-positive bugs to understand why
$\tool$ detected them.  Among them, 17 bugs are due to extraction failure
of mechanized specifications caused by wrong writing styles.
Because ECMAScript is written in HTML, $\jiset$ extracts abstract algorithms
using the \code{emu-alg} HTML tag.  Unfortunately, several abstract algorithms are
defined with the opening tag \code{<emu-alg>} but without the closing tag \code{</emu-alg>},
which causes extraction failure of mechanized specifications leading
to false-positive bugs.  The remaining 47 bugs are due to imprecise analysis.
We found that 28 bugs are due to imprecise analysis of the
conditions of assertions and branches for specific function calls.
For example, consider the following algorithm step for \textbf{GetValue}:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \includegraphics[width=0.7\columnwidth]{img/adv-refine-example}
  \vspace*{-0.5em}
\end{figure} \noindent
Since \textbf{IsPropertyReference} always returns
$\code{false}$ when the $\code{Base}$ field of a given reference record is
$\nconst{unresolvable}$, the field access $V$.[[Base]] cannot be
$\nconst{unresolvable}$ on line 4.a.  However, because the type analysis
does not compute such information,
$\nconst{unresolvable}$ is also passed as the argument of \textbf{ToObject}.
We believe that an advanced refinement technique can
resolve this problem by pruning out infeasible field types depending on specific contexts.

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.24\textwidth}
    \includegraphics[width=\textwidth]{img/compare-iter}
    \caption{The histogram of iterations}
  \end{subfigure}
  \begin{subfigure}[b]{0.24\textwidth}
    \includegraphics[width=\textwidth]{img/ratio-iter}
    \caption{The ratio of iterations}
  \end{subfigure}
  \begin{subfigure}[b]{0.24\textwidth}
    \includegraphics[width=\textwidth]{img/compare-time}
    \caption{The histogram of time}
  \end{subfigure}
  \begin{subfigure}[b]{0.24\textwidth}
    \includegraphics[width=\textwidth]{img/ratio-time}
    \caption{The ratio of time}
  \end{subfigure}
  \caption{Comparison of iterations and analysis time without refinement
  (\stextsf{no-refine}) and with refinement (\stextsf{refine})}
  \label{fig:performance-compare}
  \vspace*{-1.5em}
\end{figure}

\subsection{Effectiveness of Refinement}\label{sec:effect-refine}

\begin{table*}
  \centering
  \caption{Type-related specification bugs newly detected by $\tool$ in the
  official draft of ECMAScript 2021 (ES12)}
  \label{table:new-bug}
  \resizebox{\textwidth}{!}{%
  \begin{tabular}{@{}c@{~}?c|@{~}c@{~}|l|c|@{~}c@{~}|@{~}r@{}}
    \multicolumn{1}{@{}c?}{\textbf{Name}} &
    \multicolumn{1}{c}{\textbf{Feature}} &
    \multicolumn{1}{@{}c@{~}}{\textbf{\#}} &
    \multicolumn{1}{c}{\textbf{Description}} &
    \multicolumn{1}{@{~}c@{~}}{\textbf{Checker}} &
    \multicolumn{1}{@{}c}{\textbf{Created}} &
    \multicolumn{1}{@{}c@{~}}{\textbf{Life Span}}\\\specialrule{.1em}{.0em}{.0em}

    ES12-1 &
    Switch &
    3 &
    \makecell[l]{
        Variables \code{hasDuplicates} and \code{hasUndefinedLabels} are already \\
        defined in algorithms for \jscode{case} blocks of \jscode{switch}
        statements.
    } &
    \stextsf{\footnotesize Reference} &
    2015-09-22 &
    1,996 days\\\hline

    ES12-2 &
    Try &
    3 &
    \makecell[l]{
        Variables \code{hasDuplicates} and \code{hasUndefinedLabels} are already \\
        defined in algorithms for \jscode{try} statements.
    } &
    \stextsf{\footnotesize Reference} &
    2015-09-22 &
    1,996 days\\\hline

    ES12-3 &
    Arguments &
    1 &
    \makecell[l]{
        A variable \code{index} is already defined in
        \textbf{CreateMappedArgumentsObject}.
    } &
    \stextsf{\footnotesize Reference} &
    2015-09-22 &
    1,996 days\\\hline

    ES12-4 &
    Array &
    2 &
    \makecell[l]{
        A variable \code{succeeded} is already defined in algorithms for
        \jscode{Array} objects.
    } &
    \stextsf{\footnotesize Reference} &
    2015-09-22 &
    1,996 days\\\hline

    ES12-5 &
    Async &
    1 &
    \makecell[l]{
        A variable \code{value} is already defined in \textbf{Evaluation} for
        \jscode{yield} expressions.
    } &
    \stextsf{\footnotesize Reference} &
    2015-09-22 &
    1,996 days\\\hline

    ES12-6 &
    Class &
    1 &
    \makecell[l]{
        A variable \code{ClassHeritage} is not defined in \textbf{Contains}
        \\ for tails of \jscode{class} declarations.
    } &
    \stextsf{\footnotesize Reference} &
    2015-09-22 &
    1,996 days\\\hline

    ES12-7 &
    Branch &
    1 &
    \makecell[l]{
        A variable \code{Statement} is not defined in \textbf{EarlyErrors} for
        \jscode{if} statement.
    } &
    \stextsf{\footnotesize Reference} &
    2015-09-22 &
    1,996 days\\\hline

    ES12-8 &
    Arguments &
    2 &
    \makecell[l]{
        Abrupt completions are used in \textbf{DefineOwnProperty} and
        \textbf{GetOwnProperty} \\ for \jscode{arguments} objects without any
        checks.
    } &
    \stextsf{\footnotesize Operand} &
    2015-12-16 &
    1,910 days\\
  \end{tabular}
  }
  \vspace*{-1.5em}
\end{table*}

We measured the effectiveness of the condition-based refinement by comparing
the performance and the analysis precision of $\tool$
without (\stextsf{no-refine}) and with refinement (\stextsf{refine}).

For performance comparison, Figure~\ref{fig:performance-compare} presents the
iterations and analysis time without and with refinement.
Figures~\ref{fig:performance-compare}(a) and
\ref{fig:performance-compare}(c) are histograms of the iterations and analysis
time, respectively, and Figures~\ref{fig:performance-compare}(b) and
\ref{fig:performance-compare}(d) are scatter charts for their ratios.
Without refinement, the type analysis took 91.9 seconds with 261.5K
iterations on average.  After applying refinement, the number of iterations is increased
at least 0.99x, at most 1.36x, and 1.16x on average, and
the analysis time is increased at least 1.05x, at most 1.99x,
and 1.41x on average.

Table~\ref{table:precision} shows the analysis precision without refinement,
with refinement, and their difference.  The refinement improved the analysis
precision from 33.0\% to 59.2\% by removing 122
false-positive bugs and detecting one more true bug.  Among six
bug kinds, the most significant improvement is for non-numeric
operand bugs (\stextsf{NoNumber}) from 3.1\% to 33.3\%
by removing 59 false-positive bugs.  The refinement technique
successfully prunes out non-numeric values for numeric types.
The refinement significantly increased the analysis precision also for unknown
variable bugs (\stextsf{UnknownVar}) and assertion failures
(\stextsf{Assertion}) by removing 29 and 25 false-positive bugs, respectively.
Because $\tool$ can precisely analyze callees of function invocations
without refinement, we found no improvement for missing
parameter bugs (\stextsf{MissingParam}).

\subsection{Detection of New Bugs}\label{sec:new-bug}

Among 93 true bugs detected by $\tool$, 14 are newly
detected and still exist in the latest version of ECMAScript.
Table~\ref{table:new-bug} summarizes the bugs, their
related JavaScript language features, and their life spans.
Except for two bugs in ES12-8, all bugs were introduced in the initial
commit of the open development on September 22, 2015.
Thus, 12 newly detected bugs last for 1,996 days until March 9, 2021.
The two bugs in ES12-8 were created when a contributor introduced
the prefixes \textbf{?} and \textbf{!} on December 16, 2015, and they last for 1,910 days.
We reported the newly detected bugs to TC39, and all of them were
confirmed by the committee and will be fixed in ECMAScript 2022 (ES13).

ES12-1 contains three bugs due to duplicated variable declarations
in three syntax-directed algorithms for the \jscode{case} block
of the \jscode{switch} statement:
\code{hasDuplicates} in \textbf{ContainsDuplicateLabels} and
\code{hasUndefinedLabels} in \textbf{ContainsUndefinedBreakTarget} and
\textbf{ContainsUndefinedContinueTarget}.
A \jscode{case} block optionally contains \jscode{case} clauses.
In the beginning of three algorithms,
\code{hasDuplicates} or \code{hasUndefinedLabels} is defined if the clauses exist.
However, because the same variable is defined again after the conditional steps,
three algorithms for \jscode{case} blocks
with \jscode{case} clauses always have the duplicated variable declaration bugs for
\code{hasDuplicates} or \code{hasUndefinedLabels}.
Similarly, ES12-2 contains three bugs caused by the same reason in the
same abstract algorithms for the \jscode{try} statement.

The bug in ES12-3 is a reference bug for a duplicated declaration of the variable
\code{index} in the abstract algorithm \textbf{CreateMappedArgumentsObject}.
For each function call in JavaScript programs, an \jscode{arguments} object is
created by \textbf{CreateMappedArgumentsObject}.  In the algorithm, the variable
\code{index} is defined to handle the index of a given list of arguments.
However, the variable is defined twice in step 14 and 17 of the algorithm.

ES12-4 contains two reference bugs for the already defined variable \code{succeeded}
in \textbf{DefineOwnProperty} of \jscode{Array} objects and
\textbf{ArraySetLength}.  The \jscode{Array} objects are
not ordinary objects and have special algorithms for specific behaviors.
Two such algorithms are wrapper algorithms of
\textbf{OrdinaryDefineOwnProperty}, which updates object properties.
While they define the variable \code{succeeded} to represent the result of
\textbf{OrdinaryDefineOwnProperty}, the variable is defined twice in a specific condition.

The bug in ES12-5 is a reference bug for the already defined variable \code{value}
in \textbf{Evaluation} of the \jscode{yield *} expression.
  In the evaluation of \jscode{yield * e}, the variable \code{value} is defined twice to
represent 1) the evaluation result of the given expression \jscode{e} in step
3, and 2) the iterator value in step 7.c.viii.1.

The bug in ES12-6 is a reference bug for the unknown variable \code{ClassHeritage}
in \textbf{Contains} for the tails of \jscode{class} declarations.  A tail of a
\jscode{class} declaration consists of an optional class extension with the
\jscode{extends} keyword and a class body.  When the
optional class extension does not exist, the variable \code{ClassHeritage} is not
defined but the \textbf{Contains} algorithm tries to access it without any check
of its existence.

The bug in ES12-7 is a reference bug for the unknown variable \code{Statement} in
\textbf{EarlyErrors} for the \jscode{if} statement.  In syntax-directed algorithms,
when a production produces multiple sub-ASTs, it uses ordinal numbers as prefixes of variables.
Because the \jscode{if} statement contains two sub-ASTs
produced by the \textit{Statement} production, the ordinal number prefixes
are necessary for the variable \code{Statement}.  However, the
\textbf{EarlyErrors} algorithm for the \jscode{if} statement uses the variable without any
ordinal number prefixes.

ES12-8 contains two operand type bugs related to abrupt completions in
\textbf{DefineOwnProperty} and \textbf{GetOwnProperty} for \jscode{arguments} objects.
The two algorithms define or get own properties of \jscode{arguments} objects.
They use the \textbf{Get} algorithm, which returns JavaScript
values stored in object properties or abrupt completions.
Thus, they should check whether the results of \textbf{Get} are abstract completions or not
before using them but they use the results without any checking of abrupt completion.