### 2021-10-07 paper/pldi22/intro.tex
Added paper

\section{Introduction}\label{sec:intro}

\begin{itemize}
  \item \todo
\end{itemize}