### 2021-10-07 paper/pldi22/related.tex

\section{Related Work}\label{sec:related}

\begin{itemize}
  \item Semantics of programming languages: A tool-oriented approach~\cite{lda}
  \item LISA: An interactive environment for programming language
    development~\cite{lisa}
  \item When and how to develop domain-specific languages~\cite{dsl-survey}
  \item Incremental programming language development~\cite{ipld}
\end{itemize}

\todo