### 2021-10-12 paper/pldi22/abstract.tex
Writing intro

\begin{abstract}\label{sec:abstract}
  JavaScript is a widely-used programming language in various fields. Recently,
  it has rapidly evolved with a yearly release cadence and open development
  process to adapt users' demands to the language quickly. In this fast-evolving
  nature of JavaScript, the need for \textit{language design assistants} has
  grown to understand original or revised semantics correctly in the revision
  process. Unfortunately, while several tools have been proposed as language
  design assistants for JavaScript in academia and industry, they should be
  manually updated to represent revised semantics every time. Besides, they lack
  information about the relationship between specific semantics with conformance
  tests.

  To alleviate this problem, we present $\tool$, a \textit{live} language design
  assistant for JavaScript, which automatically synchronizes with a given
  language specification and shows related conformance tests in the revision
  process. Moreover, to better comprehend language semantics, we present a way
  to focus on specific semantics by automatically reducing semantics and
  filtering out conformance tests using user-defined \textit{syntactic views},
  abstract syntax trees with abstracted nodes. In this paper, we conduct several
  case studies for $\tool$ with historical semantic updates to demonstrate the
  effectiveness of live language design with syntactic views. Moreover, we
  experimentally show that syntactic views successfully reduce the
  \inred{XX.X}\% of semantics and filter out \inred{XX.X}\% of conformance tests
  on average.
\end{abstract}