### 2021-10-13 paper/pldi22/abstract.tex
Updated abstract

\begin{abstract}\label{sec:abstract}
  JavaScript is a widely-used programming language in various fields. Recently,
  it has rapidly evolved with a yearly release cadence and open development
  process to adapt users' demands to the language quickly. In this fast-evolving
  nature of JavaScript, the need for \textit{language design assistants} has
  grown to understand target semantics and to synchronize three language
  components: specifications, conformance tests, and language-based tools.
  Unfortunately, while several language design assistants for JavaScript have
  been proposed in academia and industry, they have two limitations: 1) no
  functionalities to focus the target semantics and 2) manual or slow updates of
  language components.

  To alleviate this problem, we present $\tool$, a \textit{live} language design
  assistant for JavaScript with \textit{syntactic views}. First, our tool
  provides a way to focus on target semantics using syntactic views, abstract
  syntax trees with abstracted nodes. It automatically reduces semantics and
  filters out conformance tests depending on user-defined syntactic views.
  Second, $\tool$ automatically and quickly synchronizes language components in
  the language design process. It detects which parts of semantics are revised
  and partially updates specifications and conformance tests. For evaluation, we
  demonstrate the effectiveness of $\tool$ with randomly selected syntactic
  views and randomly modified semantics. With randomly selected syntactic views,
  it reduces the \inred{XX.X}\% of semantics and filters out \inred{XX.X}\% of
  conformance tests on average. With randomly modified semantics, it shortens
  the time to synchronize three language components by \inred{XX.X}\%.
\end{abstract}