### 2021-10-13 paper/pldi22/abstract.tex
Writing intro

\begin{abstract}\label{sec:abstract}
  JavaScript is a widely-used programming language in various fields. Recently,
  it has rapidly evolved with a yearly release cadence and open development
  process to adapt users' demands to the language quickly. In this fast-evolving
  nature of JavaScript, the need for \textit{language design assistants} has
  grown to understand the current language semantics and to update it with new
  language features in a tool-based approach. Unfortunately, existing tools have
  two limitations related to the \textit{understanding} and \textit{update} of
  language semantics. They 1) do not provide any automatic ways to focus on
  target semantics and 2) manually or slowly synchronize three language
  components: specifications, conformance tests, and language-based tools.

  To alleviate this problem, we present $\tool$, a \textit{live} language design
  assistant for JavaScript with \textit{syntactic views}. First, our tool
  provides a way to enhance the understanding of the target semantics using a
  syntactic view, an abstract syntax tree containing abstracted nodes. It
  automatically reduces semantics and filters out conformance tests depending on
  user-defined syntactic views. Second, $\tool$ automatically and quickly
  synchronizes language components in the language design process. It detects
  which parts of semantics are revised and partially updates related parts in
  specifications, conformance tests, and language-based tools. For evaluation,
  we experimentally demonstrate the effectiveness of $\tool$. With randomly
  selected syntactic views, our tool reduces the \inred{XX.X}\% of semantics and
  filters out \inred{XX.X}\% of conformance tests on average. With randomly
  modified semantics, it shortens the time to synchronize three language
  components by \inred{XX.X}\% on average.
\end{abstract}