### 2021-10-20 paper/pldi22/related.tex
Added appendix for PE for IRES / More related work

\section{Related Work}\label{sec:related}

\begin{itemize}
  \item JavaScript specificaiton: ECMAScript (ES12, 2021)~\cite{es12}
  \item JavaScript tests: Test262~\cite{test262}
  \item JavaScript tools: engines~\cite{v8, jscore, chakra, spidermonkey},
    static analyzers~\cite{safe, safe2, tajs, wala, jsai},
    debugger~\cite{jsexplain}, verification tools~\cite{javert, javert2,
    ad-safety, javanni}, symbolic execution~\cite{symbolic-js, sym-js, expo-se},
    concolic testing~\cite{jalangi, type-conc-test}.
  \item Language Design Assistant (LDA)~\cite{lda, lisa, ipld, asf-sdf,
    meta-env, faustine}
  \item Domain Specific Language (DSL)~\cite{dsl-survey, dsl-survey2}
  \item Live Programming~\cite{omnicode, situ-vis, proj-box}
  \item JISET family: JISET~\cite{jiset}, JEST~\cite{jest}, JSTAR~\cite{jstar}.
  \item Partial Evaluation~\cite{peval, peval-survey}, Program
    Transformation~\cite{trans-ai}.
\end{itemize}

\todo