### 2021-11-02 paper/pldi22/formal.tex
More formalization

\section{JavaScript Semantics Specialization with Syntactic Views}\label{sec:formal}

In this section, we introduce a JavaScript semantics specialization with
syntactic views.  We first introduce $\ires$ as a specification language to
describe JavaScript semantics. To provide options to select target semantics, we
also introduce syntactic views as an abstraction of JavaScript ASTs.  Then, we
define a JavaScript semantics specialization with syntactic views as a partial
evaluation of $\ires$ programs, which is a combination of static analysis and
program transformation. Finally, we prove the semantics preservation of the
partial evaluation.

\subsection{$\ires$: Intermediate Representations for ECMAScript}

We first introduce $\ires$, an \textbf{I}ntermediate \textbf{R}epresentations
for \textbf{E}CMA\textbf{S}cript, as a specification language for JavaScript to
describe JavaScript semantics. We define its abstract syntax, states, and
concrete semantics in the remainder of this section.

\subsubsection{Abstract Syntax}

\[
  \begin{array}{lr@{~}c@{~}c@{~}c@{~}l}
    \text{Programs} & \progset &\ni& \prog &=& (\initstset{\prog},
    \getfunc{\prog}, \getinst{\prog}, \getnext{\prog})\\

    \text{Functions} & \funcset &\ni& \func &::=&
    \kwdef \; \kwrl \varx^* \kwrr \; \lab\\

    \text{Variables} & \varset &\ni& \varx\\

    \text{Instructions} & \instset &\ni& \inst &::=&
    \refer \kweq \expr \mid
    \varx \kweq \kwcl \kwcr \mid
    \varx \kweq \expr \kwrl \expr^* \kwrr \mid \\&&&&&
    \kwif \; \expr \; \lab \; \lab \mid
    \kwret \; \expr\\

    \text{Labels} & \labset &\ni& \lab\\

    \text{Expressions} & \exprset &\ni& \expr &::=&
    \pval \mid
    \op \kwrl \expr^* \kwrr \mid
    \refer\\

    \text{References} & \referset &\ni& \refer &::=&
    \varx \mid \expr \kwsl \expr \kwsr\\
  \end{array}
\]

An $\ires$ program $\prog = (\initstset{\prog}, \getfunc{\prog},
\getinst{\prog}, \getnext{\prog}) \in \progset $ consists of initial states and
three mappings; $\getfunc{\prog}: \labset \rightarrow \funcset$ and
$\getinst{\prog}: \labset \rightarrow \instset$ map labels to their functions
and instructions, respectively, and $\getnext{\prog}: \labset \rightarrow
\labset$ maps labels to their next labels, where a label $\lab \in \labset$
denotes a program point.  A function $\func \in \funcset$ is defined with its
parameters and body label.  For presentation brevity, we assume that no global
or captured variables exist in this paper.  An instruction $\inst \in \instset$
is an assignment $\refer \kweq \expr$, an object allocation $\varx \kweq \kwcl
\kwcr$, a function call $\varx \kweq \expr \kwrl \expr^* \kwrr$, a branch $\kwif
\; \expr \; \lab \; \lab$, or a return instruction $\kwret \; \expr$.  An
invocation of an abstract algorithm in ECMAScript is compiled to a function call
instruction with a new temporary variable.  We represent loops using branch
instructions with cyclic pointing of labels in $\getnext{\prog}$.  An expression
is a primitive value $\pval$, an operation $\op \kwrl \expr^* \kwrr$, or a
reference $\refer$.  A reference is a variable $\varx$ or an object field $\expr
\kwsl \expr \kwsr$.  We write $\expr.\varf$ to briefly represent $\expr \kwsl
\code{"f"} \kwsr$.

\subsubsection{States}

\[
  \begin{array}{lr@{~}c@{~}l@{~}c@{~}l}
    \text{States} & \st &\in& \stset &=&
    \labset \times \ctxtset^* \times \heapset \times \envset\\

    \text{Calling Contexts} & \ctxt &\in& \ctxtset &=&
    \labset \times \envset \times \varset\\

    \text{Heaps} & \heap &\in& \heapset &=&
    \addrset \finmap \objset\\

    \text{Objects} & \obj &\in& \objset &=&
    \strset \finmap \valset\\

    \text{Environments} & \env &\in& \envset &=&
    \varset \finmap \valset\\

    \text{Values} & \val &\in& \valset &=&
    \pvalset \uplus \addrset \uplus \treeset \uplus \funcset\\

    \text{Primitive Values} & \pval &\in& \pvalset &=&
    \boolset \uplus \intset \uplus \strset \uplus \cdots\\

    \text{JavaScript ASTs} & \tree &\in& \treeset\\
  \end{array}
\]

States $\stset$ consist of labels $\labset$, calling context stacks
$\ctxtset^*$, heaps $\heapset$, and environments $\envset$.  A calling context
$\ctxt \in \ctxtset$ consists of a label denoting the return point, a caller's
environment, and a return variable.  A heap $\heap \in \heapset$ is a finite
mapping from addresses to objects, and an object $\obj \in \objset$ is a finite
mappings from strings to values.  Each object allocation $\kwcl \kwcr$ creates a
unique address $\addr \in \addrset$ different from existing addresses.  An
environment $\env \in \envset$ is a finite mapping from variables to values. A
value $\val \in \valset$ is a primitive value $\pval \in \pvalset$ (e.g., a
boolean value $\bool \in \boolset$, an integer $k \in \intset$, or a string
$\str \in \strset$), an address $\addr \in \addrset$, a JavaScript AST $\tree
\in \treeset$, or a function $f \in \funcset$.

$\ires$ is a specification language for JavaScript that treats JavaScript ASTs
as its values. To handle JavaScript ASTs in $\ires$, we define them with tree
nodes $\nodeset$ as follows:
\[
  \begin{array}{l@{~}c@{~}l@{~}c@{~}l}
    \treeset &\ni& \tree &::=& \ty_k \langle \node^* \rangle\\
    \nodeset &\ni& \node &::=& \str \mid \tree\\
  \end{array}
\]
A JavaScript AST $\ty_k \langle \node_1, \cdots, \node_n \rangle$ denotes $k$-th
alternative in the syntactic production of nonterminal symbol $\ty$ with
multiple tree nodes $\node_1, \cdots, \node_n$.  A tree node is a string for a
terminal symbol or another tree for a nonterminal symbol. We define several
notations to easily deal with JavaScript ASTs.  The notation $\ty_k.\eval$ denotes
an \esalg{Evaluation} function of $k$-th alternative in the production $\ty$.
Similarly, the notation $\tree.\eval$ denotes the \esalg{Evaluation} function of
the AST $\tree$, and it is same with $\ty_k.\eval$ when $\tree = \ty_k \langle
\cdots \rangle$. The \esalg{Evaluation} of each AST takes the AST itself and its
tree nodes that are nonterminals as arguments.  The notation $\getsubs(\tree)$
denotes tree nodes that are subtrees of $\tree$.  For example,
Figure~\ref{fig:coalesce-prod} shows a syntactic production for coalesce
expressions.  Consider the following JavaScript coalesce expression:
\begin{lstlisting}[style=JS]
                    42 ?? true
\end{lstlisting}
Then, the following AST is produced as its parsing result:
\[
  \begin{array}{lcl}
    \tree_0 &=&
    \essyn{CoalesceExpression}_0 \langle \tree_1, \code{"??"}, \tree_2 \rangle\\

    \tree_1 &=&
    \essyn{CoalesceExpressionHead}_1 \langle \tree_3 \rangle\\

    \tree_2 &=&
    \essyn{BitwiseORExpression}_0 \langle \cdots \rangle\\

    \tree_3 &=&
    \essyn{BitwiseORExpression}_0 \langle \cdots \rangle\\
  \end{array}
\]
\begin{figure}[H]
  \includegraphics[width=.8\columnwidth]{img/ast-example.png}
\end{figure}
\noindent Its \esalg{Evaluation} function $\essyn{CoalesceExpression}_0.\eval$
takes three subtrees as arguments annotated by $\tree_0$, $\tree_1$, and
$\tree_2$ in the figure. The ASTs $\tree_0$, $\tree_1$, $\tree_2$, and $\tree_3$
are subtrees of $\tree_0$ (i.e., $\tree_0 \subtree \tree_0, \cdots, \tree_3
\subtree \tree_0$), and $\getsubs(\tree_0) = [\tree_1, \tree_2] \wedge
\getsubs(\tree_1) = [\tree_3]$.

\begin{figure}
  \centering
  \includegraphics[width=.8\columnwidth]{img/coalesce-prod.png}
  \caption{A JavaScript syntactic production for coalesce expressions}
  \label{fig:coalesce-prod}
\end{figure}

\subsubsection{Concrete Semantics}

The concrete semantics $\sem{\prog}$ of an $\ires$ program $\prog$ is defined as
follows:
\[
  \sem{\prog} = \{ \st \in \stset \mid \initst \in \initstset{\prog} \wedge
  \initst \trans{\prog}^* \st \}
\]
where $\trans{\prog}^*$ denotes one or more repetition of $\trans{\prog}$, and
$\st \trans{\prog} \st'$ if and only if $\st = (\lab, \_, \_, \_)$ and
$\sem{\getinst{\prog}(\lab)}(\st) = \st'$. Another way to represent the concrete
semantics is to define a transfer function $\transfer: \powerset{\stset}
\rightarrow \powerset{\stset}$:
\[
  \sem{\prog} = \lim_{n \rightarrow \infty}\transfer^n(\initstset{\prog})
\]
where
\begin{equation}\label{eqn:transfer}
  \transfer(S) = \{ \st' \mid \st \in S \wedge \st \trans{\prog} \st' \}
\end{equation}
Now, we define the denotational semantics of expressions, instructions, and
functions.

\paragraph{Expressions} We define the denotational semantics of expressions with
the following form:
\[
  \framebox{$\sem{\expr}: \stset \rightarrow \valset$}
\]
For each expression $\expr \in \exprset$, its semantics $\sem{\expr}$ takes a
state and returns a value as the result of expression. We define four different
cases in the semantics of expressions as
follows:
\begin{itemize}
  \item \underline{Primitive Values}:
    \[
      \sem{\pval}(\st) = \pval
    \]

  \item \underline{Operations}:
    \[
      \sem{\op \kwrl \expr_1, \cdots, \expr_n \kwrr}(\st) =
      \op(\val_1, \cdots, \val_n)
    \]
    where $\forall 1 \leq j \leq n. \; \sem{\expr_k}(\st) = \val_j$

  \item \underline{Variable Lookups}:
    \[
      \sem{\varx}(\st) = \env(\varx)
    \]
    where $\st = (\_, \_, \_, \env)$

  \item \underline{Field Lookups}:
    \[
      \sem{\expr_0 \kwsl \expr_1 \kwsr}(\st) = \val
    \]
    where
    \[
      \begin{array}{l@{~}c@{~}ll}
        \val_0 &=& \sem{\expr_0}(\st) &\wedge\\
        \val_1 &=& \sem{\expr_1}(\st) &\wedge\\
        \st &=& (\_, \_, \heap, \_) &\wedge\\
        \val &=& \left\{
          \begin{array}{ll}
            \heap(\addr)(\str)
            & \text{if} \; \val_0 = \addr \wedge \val_1 = \str\\

            \tree_j
            & \text{if} \; \val_0 = \ty_k \langle \tree_1, \cdots, \tree_n
            \rangle \wedge \val_1 = j\\

            \ty_k.\eval
            & \text{if} \; \val_0 = \ty_k \langle \tree_1, \cdots, \tree_n
            \rangle \wedge \val_1 = \code{"eval"}\\
          \end{array}
        \right.\\
      \end{array}
    \]
\end{itemize}

\paragraph{Instructions} We define the denotational semantics of instructions
with the following form:
\[
  \framebox{$\sem{\inst}: \stset \rightarrow \stset$}
\]
For each instruction $\inst \in \instset$, its semantics $\sem{\inst}$ takes a
state and returns an updated state. We define six different cases in the
semantics of instructions as follows:

\begin{itemize}
  \item \underline{Variable Assignments}:
    \[
      \sem{\varx \kweq \expr}(\st) =
      (\getnext{\prog}(\lab), \ctxts, \heap, \env[\varx \mapsto \val])
    \]
    where $\sem{\expr}(\st) = ((\lab, \ctxts, \heap, \env), \val)$

  \item \underline{Field Assignments}:
    \[
      \sem{\expr_0 \kwsl \expr_1 \kwsr \kweq \expr_2}(\st) =
      (\getnext{\prog}(\lab), \ctxts, \heap[\addr \mapsto \obj'], \env)
    \]
    where
    \[
      \begin{array}{l@{~}c@{~}ll}
        \sem{\expr_0}(\st) &=& (\st', \addr) &\wedge\\
        \sem{\expr_1}(\st') &=& (\st'', \str) &\wedge\\
        \sem{\expr_2}(\st'') &=& ((\lab, \ctxts, \heap, \env), \val) &\wedge\\
        \obj &=& \heap(\addr) &\wedge\\
        \obj' &=& \obj[\str \mapsto \val]\\
      \end{array}
    \]

  \item \underline{Object Allocations}:
    \[
      \sem{\varx \kweq \kwcl \kwcr}(\st) =
      (\lab, \ctxts, \heap[\addr \mapsto \epsilon], \env[\varx \mapsto
      \addr])
    \]
    where $\st = (\lab, \ctxts, \heap, \env) \wedge \addr \not\in
    \text{Domain}(\heap)$

  \item \underline{Function Calls}:
    \[
      \sem{\varx \kweq \expr \kwrl \expr_1 \cdots \expr_n \kwrr}(\st) =
      (\lab_\varf, \ctxt :: \ctxts, \heap, \env')
    \]
    where
    \[
      \begin{array}{l@{~}c@{~}ll}
        \sem{\expr}(\st) &=& (\st_0, \kwdef \; \kwrl \varx_1, \cdots, \varx_n
        \kwrr \; \lab_\varf) &\wedge\\
        \sem{\expr_j}(\st_{j-1}) &=& (\st_j, \val_j) \;
        [\forall 1 \leq j \leq n] &\wedge\\
        \st_n &=& (\lab, \ctxts, \heap, \env) &\wedge\\
        \env' &=& [\varx_1 \mapsto \val_1, \cdots, \varx_n \mapsto \val_n]
        &\wedge\\
        \ctxt &=& (\getnext{\prog}(\lab), \env, \varx)\\
      \end{array}
    \]

  \item \underline{Branches}:
    \[
      \sem{\kwif \; \expr \; \lab_\vart \; \lab_\varf}(\st) =
      \left\{
        \begin{array}{ll}
          (\lab_\vart, \ctxts, \heap, \env) & \text{if} \; \val = \true\\
          (\lab_\varf, \ctxts, \heap, \env) & \text{if} \; \val = \false\\
        \end{array}
      \right.
    \]
    where $\sem{\expr}(\st) = ((\lab, \ctxts, \heap, \env), \val)$

  \item \underline{Returns}:
    \[
      \sem{\kwret \; \expr}(\st) = (\lab, \ctxts, \heap, \env[\varx \mapsto
      \val])
    \]
    where $\sem{\expr}(\st) = ((\_, (\lab, \env, \varx) :: \ctxts, \heap, \_),
    \val)$
\end{itemize}

\subsection{Syntactic Views}

A syntactic view $\atree \in \atreeset$ is an abstract JavaScript AST with
abstract tree nodes $\anodeset$:
\[
  \begin{array}{l@{~}c@{~}l@{~}c@{~}l}
    \atreeset &\ni& \atree &::=& \ty_k \langle \anode^* \rangle \mid \ty\\
    \anodeset &\ni& \anode &::=& \str \mid \atree\\
  \end{array}
\]
where $\ty$ denotes all ASTs whose nonterminal symbol $\ty$. We define the
concretization of abstract trees $\atreegamma: \atreeset \rightarrow
\powerset{\treeset}$ and that of abstract nodes $\anodegamma: \anodeset
\rightarrow \powerset{\nodeset}$ as follows:
\[
  \begin{array}{lcl}
    \atreegamma(\ty_k \langle \anode_1, \cdots, \anode_n \rangle) &=&
    \{ \ty_k \langle \node_1, \cdots, \node_n \rangle \mid \node_j \in
    \anodegamma(\anode_j) \}\\

    \atreegamma(\ty) &=&
    \{ \tree \in \treeset \mid \tree = \ty_k \langle \cdots \rangle \}\\

    \anodegamma(\str) &=& \{ \str \}\\

    \anodegamma(\atree) &=& \atreegamma(\atree)\\
  \end{array}
\]
% \[
%   \begin{array}{lcl}
%     \anode \order \anode' &\Leftrightarrow& \left\{
%       \begin{array}{ll}
%         \anode = \anode' &\vee\\
% 
%         \anode' = \ty &\vee\\
% 
%         (\anode = \atree \wedge \anode' = \atree' \wedge \atree \order
%         \atree')\\
%       \end{array}
%     \right.\\
% 
%     \atree \order \atree' &\Leftrightarrow& \left\{
%       \begin{array}{ll}
%         \atree = \atree' &\vee\\
% 
%         \atree' = \treetop &\vee\\
% 
%         \begin{array}{@{}l@{}}
%           \atree = \ty_k \langle \atree_1, \cdots, \atree_n \rangle \wedge
%           \atree' = \ty_k \langle \atree'_1, \cdots, \atree'_n \rangle \wedge\\
%           \forall j. \; \tree_j \order \tree'_j\\
%         \end{array}\\
%       \end{array}
%     \right.\\
% 
%     \atree \join \atree' &=& \left\{
%       \begin{array}{ll}
%         \atree' & \text{if} \; \atree \order \atree'\\
% 
%         \atree & \text{if} \; \atree \revorder \atree'\\
% 
%         \ty_k \langle \atree''_1, \cdots, \atree''_n \rangle &
%         \text{if} \; \atree = \ty_k \langle \atree_1, \cdots, \atree_n \rangle
%         \wedge\\&
%         \phantom{\text{if} \;} \atree = \ty_k \langle \atree_1, \cdots, \atree_n
%         \rangle \wedge\\&
%         \phantom{\text{if} \;} \forall j. \; \atree''_j = \atree_j
%         \join \atree'_j\\
%         \treetop & \text{otherwise}\\
%       \end{array}
%     \right.\\
%   \end{array}
% \]
Moreover, we define abstract subtree relation $\asubtree \subseteq \atreeset
\times \treeset$:
\[
  \atree \; \asubtree \; \tree' \iff \exists \tree \in \atreegamma(\atree). \;
  \tree \subtree \tree'
\]
to filter out JavaScript programs not related to a given syntactic view.
Similar to $\getsubs(\tree)$, an abstract helper function $\agetsubs(\atree)$
denotes abstract tree nodes that are syntactic views when $\atree = \ty_k
\langle \cdots \rangle$.

\subsection{Partial Evaluation}

We define a \textit{partial evaluation} $\peval{\prog}: \atreeset \rightarrow
\progset$ of $\ires$ programs to specialize JavaScript semantics with a given
syntactic view:
\[
  \peval{\prog}(\atree) = \transform{\prog} \circ \asem{\prog} (\atree)
\]
It takes a syntactic view $\atree \in \atreeset$ to restrict arguments of the
corresponding \esalg{Evaluation} function. Then, it performs a static analysis
$\asem{\prog}$ of a given program $\prog$ with the syntactic view $\atree$, and
transforms the program to a specialized program using $\transform{\prog}$ with
the analysis result $\asem{\prog}(\atree)$.  Since the \esalg{Evaluation} of
$\ty$ might one or more $\ires$ functions, we only define the partial evaluation
when $\atree = \ty_k \langle \cdots \rangle$.  Now, we define the static
analysis and program transformation of $\ires$ programs and prove the semantics
preservation of the partial evaluation consisting of them.

\subsubsection{Static Analysis}

The static analysis $\asem{\prog}(\atree)$ performs an abstract
interpretation~\cite{ai1977, ai1992} to abstract concrete semantics under the
given given syntactic view $\atree$.  We first define abstract domains of
states, call edges, environments, and values as follows:
\begin{itemize}
  \item \underline{Abstract States}: $\aelemset$
    \[
      \begin{array}{lcl}
        \aelemset &=& \adcallset \times \ascallset \times \afsenvset\\

        \aelemgamma &:& \atreeset \rightarrow \aelemset \rightarrow
        \powerset{\stset}\\

        \aelemgamma(\atree)(\aelem) &=&
        (\adcallgamma(\atree)(\adcall) \setminus C) \uplus
        (C \cap \afsenvgamma(\afsenv))\\
        \multicolumn{3}{r}{\text{where} \; C = \ascallgamma(\atree)(\ascall)}\\

        \aelem \order \aelem' &\Leftrightarrow&
        \adcall \order \adcall' \wedge
        \ascall \order \ascall' \wedge
        \afsenv \order \afsenv'\\

        \aelem \join \aelem' &=&
        (\adcall \join \adcall', \ascall \join \ascall', \afsenv \join \afsenv')\\
      \end{array}
    \]
    where $\atree = \ty_k \langle \cdots \rangle$, $\aelem = (\adcall, \ascall,
    \afsenv)$ and $\aelem' = (\adcall', \ascall', \afsenv')$

  \item \underline{Dynamic Call Sets}: $\adcallset$
    \[
      \begin{array}{lcl}
        \adcallset &=& \powerset{\labset}\\

        \adcallgamma &:& \atreeset \rightarrow \adcallset \rightarrow
        \powerset{\stset}\\

        \adcallgamma(\atree)(\adcall) &=& \{ \st \in \stset \mid
          \st = (\lab, [\ctxt_1, \cdots, \ctxt_n], \_, \_) \wedge\\&&

          \phantom{\{ \st \in \stset \mid}
          \exists 1 \leq j \leq n. \; \ctxt_j = (\lab_j, \_, \_) \wedge\\&&

          \phantom{\{ \st \in \stset \mid}
          \lab_j \in \adcall
        \}\\

        \adcall \order \adcall' &\Leftrightarrow&
        \adcall \subseteq \adcall\\

        \adcall \join \adcall' &=&
        \adcall \cup \adcall\\
      \end{array}
    \]
    where $\atree = \ty_k \langle \cdots \rangle$

  \item \underline{Static Call Sets}: $\ascallset$
    \[
      \begin{array}{lcl}
        \ascallset &=& \powerset{\labset \times \funcset}\\

        \ascallgamma &:& \atreeset \rightarrow \ascallset \rightarrow
        \powerset{\stset}\\

        \ascallgamma(\atree)(\ascall) &=& \{ \st \in \stset \mid
          \st = (\lab_0, [\ctxt_1, \cdots, \ctxt_n], \_, \_) \wedge\\&&

          \phantom{\{ \st \in \stset \mid}
          \forall 1 \leq j \leq n. \; \ctxt_j = (\lab_j, \_, \_) \wedge\\&&

          \phantom{\{ \st \in \stset \mid}
          \exists 0 \leq m \leq n. \; \getfunc{\prog}(\lab_m) = \ty_k.\eval
          \wedge\\&&

          \phantom{\{ \st \in \stset \mid}
          \forall 1 \leq j \leq m. \; (\lab_j, \getfunc{\prog}{\lab_{j-1}}) \in
          \ascall
        \}\\

        \ascall \order \ascall' &\Leftrightarrow& \ascall \subseteq \ascall\\

        \ascall \join \ascall' &=& \ascall \cup \ascall'\\
      \end{array}
    \]
    where $\atree = \ty_k \langle \cdots \rangle$

  \item \underline{Flow-Sensitive Abstract Environments}: $\afsenvset$
    \[
      \begin{array}{lcl}
        \afsenvset &=& \labset \rightarrow \aenvset\\

        \afsenvgamma &:& \afsenvset \rightarrow \powerset{\stset}\\

        \afsenvgamma(\afsenv) &=& \{ \st \in \stset \mid
          \st = (\lab, \_, \_, \_) \wedge \st \in \aenvgamma(\afsenv(\lab))
        \}\\

        \afsenv \order \afsenv' &\Leftrightarrow&
        \forall \lab \in \labset. \; \afsenv(\lab) \order \afsenv'(\lab)\\

        \afsenv \join \afsenv' &=&
        \lambda \lab \in \labset. \; \afsenv(\lab) \join \afsenv'(\lab)\\
      \end{array}
    \]

  \item \underline{Abstract Environments}: $\aenvset$
    \[
      \begin{array}{lcl}
        \aenvset &=& \varset \rightarrow \aenvset\\

        \aenvgamma &:& \aenvset \rightarrow \powerset{\stset}\\

        \aenvgamma(\aenv) &=& \{ \st \in \stset \mid
          \forall \varx \in \varset. \;
          \sem{\varx}(\st) \in \avalgamma(\aenv(\varx))
        \}\\

        \aenv \order \aenv' &\Leftrightarrow&
        \forall \varx \in \varset. \;
        \aenv(\varx) \order \aenv'(\varx)\\

        \aenv \join \aenv' &=&
        \lambda \varx \in \varset. \;
        \aenv(\varx) \join \aenv'(\varx)\\
      \end{array}
    \]

  \item \underline{Abstract Values}: $\avalset$
    \[
      \begin{array}{lcl}
        \avalset &=& \valset \uplus \atreeset \uplus \{ \top, \bot,  \}\\

        \avalgamma &:& \avalset \rightarrow \powerset{\valset}\\

        \avalgamma(\aval) &=& \left\{
          \begin{array}{ll}
            \{ \val \} & \text{if} \; \aval = \val\\
            \atreegamma(\atree) & \text{if} \; \aval = \atree\\
            \valset & \text{if} \; \aval = \top\\
            \varnothing & \text{if} \; \aval = \bot\\
          \end{array}
        \right.\\

        \aval \order \aval' &\Leftrightarrow& \left\{
          \begin{array}{ll}
            \aval = \bot \vee \aval' = \top \vee \aval = \aval' & \vee\\
            \aval = \atree \wedge \aval' = \atree' \wedge
            \atree \order \atree' &\vee\\
            \aval = \tree \wedge \aval' = \atree' \wedge \tree \in
            \atreegamma(\atree')\\
          \end{array}
        \right.\\

        \aval \join \aval' &=& \left\{
          \begin{array}{ll}
            \aval' & \text{if} \; \aval \order \aval'\\
            \aval & \text{if} \; \aval \revorder \aval'\\
            \top & \text{otherwise}\\
          \end{array}
        \right.\\
      \end{array}
    \]
\end{itemize}

An abstract state $\aelem \in \aelemset$ is a triple of a dynamic call set, a
static call set, and a flow-sensitive abstract environment.  A dynamic call set
$\adcall \in \adcallset$ is a set of call labels representing function calls
with multiple functions or no static arguments. A static call set $\ascall \in
\ascallset$ is a set of possible static call edges from labels to function, and
it represent a set of \textit{static calling contexts} $\ascallgamma(\ascall)$
used as the scope of program transformations.  To more precisely in the static
calling contexts, we define a flow-sensitive abstract environment $\afsenv \in
\afsenvset$. It is a mapping from labels to abstract environments to represents
the shape of environments in each label. An abstract environment $\aenv \in
\aenvset$ is a mapping from variables to abstract values to describe values
stored in variables. An abstract value $\aval \in \avalset$ is a static value
$\val$, a syntactic view $\atree$, a dynamic value $\top$, or nothing $\bot$.
Now, we define abstract semantics of expressions, instructions, and programs.

\paragraph{Expressions} We first define abstract semantics of expressions as
follows:
\[
  \framebox{$\asem{\expr}: \aenvset \rightarrow \avalset$}
\]
\begin{itemize}
  \item \underline{Primitive Values}:
    \[
      \asem{\pval}(\aenv) = \pval
    \]
  \item \underline{Operations}:
    \[
      \asem{\op \kwrl \expr_1, \cdots, \expr_n \kwrr}(\aenv) = \aval
    \]
    where
    \[
      \begin{array}{l@{~}c@{~}l@{~}l}
        \asem{\expr_j}(\aenv) &=& \aval_j \; [\forall 1 \leq j \leq n]
        &\wedge\\
        \aval &=& \left\{
          \begin{array}{ll}
            \bot & \text{if} \; \exists j. \; \aval_j = \bot\\

            \op(\val_1, \cdots, \val_n) &
            \text{if} \; \forall j. \; \aval_j = \val_j\\

            \top & \text{otherwise}\\
          \end{array}
        \right.\\
      \end{array}
    \]
  \item \underline{Variable Lookups}:
    \[
      \asem{\varx}(\aenv) = \aenv(\varx)
    \]
  \item \underline{Field Lookups}:
    \[
      \asem{\expr_0 \kwsl \expr_1 \kwsr}(\aenv) = \aval
    \]
    where
    \[
      \begin{array}{lcl}
        \asem{\expr_0}(\aenv) &=& \aval_0\\
        \asem{\expr_1}(\aenv) &=& \aval_1\\
        \aval &=& \left\{
          \begin{array}{ll}
            \atree_j & \text{if} \;
            \aval_0 = \ty_k \langle \atree_1, \cdots, \atree_n \rangle \wedge
            \aval_1 = j\\

            \ty_k.\eval & \text{if} \;
            \aval_0 = \ty_k \langle \cdots \rangle \wedge
            \aval_1 = \code{"eval"}\\

            \top & \text{otherwise}\\
          \end{array}
        \right.\\
      \end{array}
    \]
\end{itemize}

\paragraph{Instructions} Then, we define the abstract semantics of instructions
as follows:
\[
  \framebox{$\asem{\inst}: \labset \times \aelemset \rightarrow \aelemset$}
\]
\begin{itemize}
  \item \underline{Variable Assignments}:
    \[
      \asem{\varx \kweq \expr}(\lab, \aelem) =
      (\adcall, \ascall, [\getnext{\prog}(\lab) \mapsto \aenv[\varx
      \mapsto \aval]])
    \]
    where
    \[
      \begin{array}{lcl}
        \aelem &=& (\adcall, \ascall, \afsenv)\\
        \aenv &=& \afsenv(\lab)\\
        \aval &=& \asem{\expr}(\aenv)\\
      \end{array}
    \]

  \item \underline{Field Assignments}:
    \[
      \asem{\expr_0 \kwsl \expr_1 \kwsr \kweq \expr_2}
      (\lab, \aelem) =
      (\adcall, \ascall, [\getnext{\prog}(\lab) \mapsto \aenv])
    \]
    where
    \[
      \begin{array}{lcl}
        \aelem &=& (\adcall, \ascall, \afsenv)\\
        \aenv &=& \afsenv(\lab)\\
      \end{array}
    \]

  \item \underline{Object Allocations}:
    \[
      \asem{\varx \kweq \kwcl \kwcr}(\lab, \aelem) =
      (\adcall, \ascall, [\getnext{\prog}(\lab) \mapsto \aenv[\varx
      \mapsto \top]])
    \]
    where
    \[
      \begin{array}{lcl}
        \aelem &=& (\adcall, \ascall, \afsenv)\\
        \aenv &=& \afsenv(\lab)\\
      \end{array}
    \]

  \item \underline{Function Calls}:
    \[
      \asem{\varx \kweq \expr \kwrl \expr_1 \cdots \expr_n \kwrr}(\lab, \aelem)
      = (\adcall', \ascall', \afsenv')
    \]
    where
    \[
      \begin{array}{l@{~}c@{~}l}
        \aelem &=& (\adcall, \ascall, \afsenv)\\
        \aenv &=& \afsenv(\lab)\\
        \aval &=& \asem{\expr}(\aenv)\\
        \aval_j &=& \asem{\expr_j}(\aenv) \; [\forall 1 \leq j \leq n]\\
        (\adcall', \ascall') &=& \left\{
          \begin{array}{ll}
            (\adcall, \ascall \cup \{ (\lab, \func) \} &
            \text{if} \; \aval = \func \wedge \exists j. \; \aval_j \neq \top\\

            (\adcall \cup \{ \lab \}, \ascall) &
            \text{otherwise}
          \end{array}
        \right.\\
        \afsenv' &=& [ \lab_\func \mapsto \aenv_\func \mid
        (\lab, \func) \in \ascall' \wedge\\&&

        \phantom{[\lab_\func \mapsto \aenv_\func \mid \;}
        \func = \kwdef \; \kwrl \varx_1, \cdots, \varx_n \kwrr
        \; \lab_\func \wedge\\&&

        \phantom{[\lab_\func \mapsto \aenv_\func \mid \;}
        \aenv_\func = [\varx_1 \mapsto \aval_1, \cdots,
        \varx_n \mapsto \aval_n] ]\\
      \end{array}
    \]

  \item \underline{Branches}:
    \[
      \asem{\kwif \; \expr \; \lab_\vart \; \lab_\varf}(\lab, \aelem) =
      (\adcall, \ascall, \afsenv'')
    \]
    where
    \[
      \begin{array}{lcl}
        \aelem &=& (\adcall, \ascall, \afsenv)\\
        \aenv &=& \afsenv(\lab)\\
        \aval &=& \asem{\expr}(\aenv)\\

        \afsenv' &=& \left\{
          \begin{array}{ll}
            [\lab_\vart \mapsto \aenv] & \text{if} \; \true \in
            \avalgamma(\aval)\\
            \epsilon & \text{otherwise}\\
          \end{array}
        \right.\\

        \afsenv'' &=& \left\{
          \begin{array}{ll}
            \afsenv'[\lab_\varf \mapsto \aenv] & \text{if} \; \false \in
            \avalgamma(\aval)\\
            \afsenv' & \text{otherwise}\\
          \end{array}
        \right.\\
      \end{array}
    \]

  \item \underline{Returns}:
    \[
      \asem{\kwret \; \expr}(\lab_0, \aelem) =
      (\adcall, \ascall, \afsenv')
    \]
    where
    \[
      \begin{array}{lcl}
        \aelem &=& (\adcall, \ascall, \afsenv)\\
        \aenv &=& \afsenv(\lab)\\
        \aval &=& \asem{\expr'}(\aenv)\\

        \func &=& \getfunc{\prog}(\lab_0, \aelem)\\
        R &=& \{ (\getnext{\prog}(\lab), \varx) \mid
          (\lab, \func) \in \ascall \wedge\\&&

          \phantom{\{ (\getnext{\prog}(\lab), \varx) \mid}
          \getinst{\prog}(\lab) = \varx \kweq \expr^\lab (\expr^\lab_1 \cdots
          \expr^\lab_n)
        \}\\

        \afsenv' &=& \lambda \lab \in \labset. \; \left\{
          \begin{array}{ll}
            [\varx \mapsto \aval]
            & \text{if} \; (\lab, \varx) \in R\\

            \epsilon
            & \text{otherwise}\\
          \end{array}
        \right.\\
      \end{array}
    \]
\end{itemize}

\paragraph{Programs} Finally, we define the abstract semantics $\asem{\prog}:
\atreeset \rightarrow \aelemset$ of a program $\prog$ with a given syntactic
view:
\[
  \asem{\prog}(\atree) = \lim_{n \rightarrow \infty}(\atransfer)^n \circ
  \entryaelem(\atree)
\]
It is the least fixed point of the abstract transfer function $\atransfer$ with
an abstract entry state $\entryaelem$ depending on the given syntactic view
$\atree$:
\begin{itemize}
  \item \underline{Abstract Transfer Function}: $\atransfer: \aelemset
    \rightarrow \aelemset$
    \[
      \atransfer(\aelem) = \aelem \join \sem{\adcall} \wedge
      \bigjoin_{\lab \in \labset}{\asem{\getinst{\prog}(\lab)}(\lab, \aelem)}
    \]
    where $\aelem = (\adcall, \ascall, \afsenv)$ and $\sem{\adcall} = (\adcall,
    \ascall, [\lab \mapsto \aenv_\lab \mid \lab \in \adcall \wedge
    \getinst{\prog}(\lab) = \varx = \expr(\cdots) \wedge \aenv_\lab = [\varx
    \mapsto \top]])$.
  \item \underline{Abstract Entry States}: $\entryaelem:
    \atreeset \rightarrow \aelemset$
    \[
      \entryaelem(\atree) = (\varnothing, \varnothing, \afsenv)
    \]
    where \[
      \begin{array}{lcl}
        \atree &=& \ty_k \langle \cdots \rangle\\

        \agetsubs(\atree) &=& [\atree_1, \cdots, \atree_n]\\

        \ty_k.\eval &=& \lambda \kwdef \; \kwrl \varx_1, \cdots, \varx_n \kwrr
        \; \entrylab\\

        \aenv &=& [\varx_1 \mapsto \atree_1, \cdots, \varx_n \mapsto
        \atree_n]\\

        \afsenv &=& [\entrylab \mapsto \aenv]\\
      \end{array}
    \]
\end{itemize}

\subsubsection{Transformations}

Using the analysis result, we define transformations for expressions,
instructions, and programs.

\paragraph{Expressions} We first define the transformation of expressions as
follows:

\[
  \framebox{$\transform{\expr}: \aenvset \rightarrow \exprset$}
\]
\[
  \transform{\expr}(\aenv) = \left\{
    \begin{array}{ll}
      \pval & \text{if} \; \asem{\expr}(\aenv) = \pval\\
      \expr & \text{otherwise}\\
    \end{array}
  \right.
\]

\paragraph{Instructions} Then, we define the transformation of instructions
as follows:
\[
  \framebox{$\transform{\inst}: \aenvset \rightarrow \instset$}
\]
\begin{itemize}
  \item \underline{Variable Assignments}:
    \[
      \transform{\varx \kweq \expr}(\aenv) =
      \varx \kweq \expr'
    \]
    where $\expr' = \transform{\expr}(\aenv)$

  \item \underline{Field Assignments}:
    \[
      \transform{\expr_0 \kwsl \expr_1 \kwsr \kweq \expr_2}(\aenv) =
      \expr'_0 \kwsl \expr'_1 \kwsr \kweq \expr'_2
    \]
    where
    \[
      \begin{array}{l@{~}c@{~}ll}
        \expr'_0 = \transform{\expr_0}(\aenv) &\wedge\\
        \expr'_1 = \transform{\expr_1}(\aenv) &\wedge\\
        \expr'_2 = \transform{\expr_2}(\aenv)\\
      \end{array}
    \]
  \item \underline{Object Allocations}:
    \[
      \transform{\varx \kweq \kwcl \kwcr}(\aenv) =
      \varx \kweq \kwcl \kwcr
    \]

  \item \underline{Function Calls}:
    \[
      \transform{\varx \kweq \expr \kwrl \expr_1 \cdots \expr_n \kwrr}(\aenv) =
      \varx \kweq \expr' \kwrl \expr'_1 \cdots \expr'_n \kwrr
    \]
    where
    \[
      \begin{array}{l@{~}c@{~}ll}
        \expr' = \transform{\expr}(\aenv) &\wedge\\
        \expr'_k = \transform{\expr_k}(\aenv) \; [\forall 1 \leq k \leq
        n]\\
      \end{array}
    \]

  \item \underline{Branches}:
    \[
      \transform{\kwif \; \expr \; \lab_\vart \; \lab_\varf}(\aenv) =
      \kwif \; \expr' \; \lab_\vart \; \lab_\varf
    \]
    where $\expr' = \transform{\expr}(\aenv)$

  \item \underline{Returns}:
    \[
      \transform{\kwret \; \expr}(\aenv) =
      \kwret \; \expr'
    \]
    where $\expr' = \transform{\expr}(\aenv)$
\end{itemize}

\paragraph{Programs} Finally, we define the program transformation as follows:
\[
  \framebox{$\transform{\prog}: \aelemset \rightarrow \progset$}
\]
\[
  \transform{\prog}(\adcall, \ascall, \afsenv) = (\initstset{\prog},
  \getfunc{\prog}, \getinst{\prog'}, \getnext{\prog})
\]
where
\[
  \begin{array}{lcl}
    \prog &=&
    (\initstset{\prog}, \getfunc{\prog}, \getinst{\prog}, \getnext{\prog})\\

    \getinst{\prog'} &=&
    \getinst{\prog}[\lab \mapsto \inst_\lab \mid
    \lab \in \labset \wedge \aenv = \afsenv(\lab) \wedge\\&&

    \phantom{\getinst{\prog}[\lab \mapsto \inst_\lab \mid}
    \inst_\lab = \transform{\getinst{\prog}(\lab)}(\aenv)]\\
  \end{array}
\]

\subsubsection{Cleanup Algorithm}

\todo

\begin{itemize}
  \item perform a syntactic def-use analysis
  \item remove unnecessary variable assignments
  \item remove unreachable instructions
\end{itemize}

\subsubsection{Semantics Preservation}

To prove the semantics preservation of the partial evaluation for each syntactic
view $\atree$, we first define a restricted semantics depending on the given
syntactic view. Then, we prove the soundness of the static analysis
$\asem{\prog}$ of the program. Finally, we prove the semantics preservation of
the program transformation $\transform{\prog}$.

\paragraph{Restricted Semantics} We first define a \textit{restricted semantics}
$\rsem{\prog}: \atreeset \rightarrow \powerset{\stset}$ with a syntactic view:
\[
  \rsem{\prog}(\atree) = \lim_{n \rightarrow \infty}\transfer^n \circ
  \entrystset(\atree)
\]
It is the least fixed point of a transfer function $\transfer: \powerset{\stset}
\rightarrow \powerset{\stset}$ defined in Equation~\ref{eqn:transfer} with the
following entry states $\entrystset: \atreeset \rightarrow \powerset{\stset}$
depending on the given syntactic view $\atree$:
\[
  \begin{array}{lcl}
    \entrystset &:& \atreeset \rightarrow \powerset{\stset}\\

    \entrystset(\atree) &=& \{ \st \in \stset \mid
      \st = (\entrylab, \_, \_, \env) \wedge\\&&

      \phantom{\{ \st \in \stset \mid}
        \tree \in \atreegamma(\atree) \wedge \forall 1 \leq j \leq n. \;
      \tree_j \in \atreegamma(\atree_j) \wedge\\&&

      \phantom{\{ \st \in \stset \mid}
        \env = [\varx_0 \mapsto \tree, \varx_1 \mapsto \tree_1, \cdots,
        \varx_n \mapsto \tree_n]
      \}\\
    \end{array}
  \]
where
\[
  \begin{array}{lcl}
    \atree &=& \ty_k \langle \cdots \rangle\\
    \ty_k.\eval &=& \kwdef \; \kwrl \varx_0, \cdots, \varx_n \kwrr \entrylab\\
    \agetsubs(\atree) &=& [\atree_1, \cdots, \atree_n]\\
  \end{array}
\]

\paragraph{Proof of Soundness} We formally define the soundness of the static
analysis $\asem{\prog}$ in Theorem~\ref{thm:sound-prog} and prove it using
several lemmas.

\begin{theorem}[Soundness of $\asem{\prog}$]\label{thm:sound-prog}
  For a given syntactic view $\atree \in \atreeset$, the static analysis
  $\asem{\prog}(\atree)$ is sound by satisfying the following condition:
  \[
    \rsem{\prog}(\atree) \subseteq \aelemgamma \circ \asem{\prog}(\atree)
  \]
\end{theorem}
\begin{proof}
  \todo
\end{proof}

\begin{lemma}[Soundness of $\entryaelem$]\label{thm:sound-entryaelem}
  For a given syntactic view $\atree \in \atreeset$, the abstract entry state
  $\entryaelem(\atree)$ is a sound approximation of the entry states
  $\entrystset(\atree)$ by satisfying the following condition:
  \[
    \entrystset(\atree) \subseteq \aelemgamma \circ \entryaelem(\atree) 
  \]
\end{lemma}
\begin{proof}
  \todo
\end{proof}

\begin{lemma}[Soundness of $\atransfer$]\label{thm:sound-atransfer}
  The abstract transfer function $\atransfer$ is a sound approximation of the
  transfer function $\transfer$ by satisfying the following condition:
  \[
    \forall \aelem \in \aelemset. \;
    \transfer \circ \aelemgamma(\aelem) \subseteq
    \aelemgamma \circ \atransfer(\aelem)
  \]
\end{lemma}
\begin{proof}
  \todo
\end{proof}

\begin{lemma}[Soundness of $\asem{\inst}$]\label{thm:sound-inst}
  For a given label $\lab \in \labset$ and its instruction $\inst =
  \getinst{\prog}(\lab)$, the abstract semantics $\asem{\inst}$ of the
  instruction is a sound approximation of its concrete semantics $\sem{\inst}$:
  \[
    \forall \aelem \in \aelemset. \;
    \st \in \aelemgamma(\aelem) \wedge \st = (\lab, \_, \_, \_) \Rightarrow
    \sem{\inst}(\st) \in \aelemgamma \circ \asem{\inst}(\lab, \aelem)
  \]
\end{lemma}
\begin{proof}
  \todo
\end{proof}

\begin{lemma}[Soundness of $\asem{\expr}$]\label{thm:sound-inst}
  The abstract semantics $\asem{\inst}$ of an expression $\expr$ is a sound
  approximation of its concrete semantics $\sem{\expr}$:
  \[
    \forall \aenv \in \aenvset. \;
    \st \in \aenvgamma(\aenv) \Rightarrow
    \sem{\expr}(\st) \in \avalgamma \circ \asem{\expr}(\aenv)
  \]
\end{lemma}
\begin{proof}
  \todo
\end{proof}

\paragraph{Proof of Semantics Preservation} Finally, we prove the semantics
prservation of the program transformation $\transform{\prog}$ in
Theorem~\ref{thm:preserve-prog} with several lemmas.

\begin{theorem}[Semantics Preservation of $\transform{\prog}$]
  \label{thm:preserve-prog}
  For a given syntactic view $\atree \in \atreeset$, a transformed program
  $\prog' = \transform{\prog}(\aelem)$ with an abstract state $\aelem =
  (\adcall, \ascall, \afsenv) \in \aelemset$ preserves the restricted semantics
  $\rsem{\prog}(\atree)$ under the static calling contexts
  $\ascallgamma(\ascall)$ if $\aelem$ is sound:
  \[
    \todo
    % \rsem{\prog}(\atree) \subseteq \aelemgamma(\aelem) \Rightarrow
    % \forall \st \in \rsem{\prog}(\atree) \cap \ascallgamma(\ascall). \;
    % (\st \trans{\prog} \st' \wedge \st \trans{\prog'} \st'') \Rightarrow (\st' =
    % \st'')
  \]
\end{theorem}
\begin{proof}
  \todo
\end{proof}