### 2021-11-04 paper/pldi22/abstract.tex
Revised abstract

\begin{abstract}\label{sec:abstract}
  JavaScript is a widely-used programming language in various fields. Recently,
  it has rapidly evolved with a yearly release cadence and open development
  process to adapt users' demands to the language quickly. In this fast-evolving
  nature of JavaScript, developers and even language designers often suffer from
  complex semantics in newly introduced language features and want to
  \textit{understand} them correctly.  To meet the demand, researchers have
  proposed several tools that enhance the understanding of JavaScript semantics.
  However, existing tools 1) require manual updates for new specifications or 2)
  do not provide additional information when users desire to understand specific
  parts of semantics.

  To alleviate this problem, we present $\tool$, a JavaScript interactive
  specification. First, it provides step-by-step program execution automatically
  synchronized with a given version of JavaScript specification. Second, our
  tool offers two different kinds of additional information for target semantics
  selected by users: 1) reduced semantics for a selected \textit{syntactic view}
  using \textit{partial evaluation} and 2) a small-sized program related to a
  selected algorithm step using \textit{delta debugging} with a given set of
  programs. For evaluation, we experimentally demonstrate the effectiveness of
  $\tool$ with the latest ECMAScript (ES12, 2021), \inred{177} syntactic views,
  and \inred{19,839} programs. Our tool successfully reduces \inred{18.9}\%
  algorithm steps for each syntactic view on average. For each algorithm step,
  it provides \inred{2.3} lines (\inred{152.3} bytes) of a program with
  automatically inferred assertions on average.
\end{abstract}