### 2021-11-04 paper/pldi22/abstract.tex
Revised intro

\begin{abstract}\label{sec:abstract}
  JavaScript is a widely-used programming language in various fields.
  Unfortunately, developers and even language designers often suffer from
  complex semantics in JavaScript and desire to \textit{understand} them
  correctly. Therefore, researchers have presented several tools for a better
  understanding of JavaScript semantics. However, existing tools 1) require
  manual updates for new specifications while JavaScript has rapidly evolved
  with various language features and 2) do not support user interactions to
  select target semantics to get additional information.

  To alleviate this problem, we present $\tool$, a JavaScript interactive
  specification. First, it provides step-by-step program execution automatically
  synchronized with a given version of JavaScript specification. Second, our
  tool supports two different types of user interactions to select target
  semantics: 1) a \textit{syntactic view selection} to reduce abstract
  algorithms using \textit{partial evaluation} and 2) an \textit{algorithm step
  selection} to get a small-sized JavaScript program related to the step using
  \textit{delta debugging} with a given set of programs. For evaluation, we
  experimentally demonstrate the effectiveness of $\tool$ with the latest
  ECMAScript (ES12, 2021), \inred{177} syntactic views, and \inred{19,839}
  programs. Our tool successfully supports step-by-step program execution in
  ES12. It reduces \inred{18.9}\% algorithm steps for each syntactic view and
  provides \inred{2.3} lines (\inred{152.3} bytes) of a program for each step on
  average.
\end{abstract}