### 2021-11-04 paper/pldi22/intro.tex

\section{Introduction}\label{sec:intro}

In the beginning, JavaScript was a scripting language designed in a ten-day
hack. However, it has become a de facto Web standard and eventually becomes one
of the dominating programming languages in various fields. For example,
Node.js\footnote{https://nodejs.org/} introduced full-stack JavaScript by
supporting server-side programming, and JavaScript has recently begun to be used
even in embedded systems and microcontrollers for the Internet of
Things\footnote{https://www.espruino.com/}. Furthermore, according to the annual
report of GitHub\footnote{\url{https://octoverse.github.com/}}, JavaScript has
consistently been the most popular programming language based on the number of
contributors to GitHub projects.

Unfortunately, developers and even language designers often suffer from complex
semantics in new language features and desire to understand them correctly.
ECMAScript is the standard specification for JavaScript maintained by Ecma
Technical Committee 39 (TC39). It describes the language semantics in abstract
algorithms consisting of English sentences. However, it is infeasible to know
all the detailed complex semantics by reading the specification even to language
designers. For example, consider a JavaScript logical nullish assignment
\jscode{x ??= y} newly introduced in the latest ECMAScript (ES12,
2021)~\cite{es12}. Its basic semantics is to assign \jscode{y} to \jscode{x}
only if \jscode{x} is nullish (\jscode{undefined} or \jscode{null}). Therefore,
it seems to have exactly same semantics with an assignment \jscode{x = x ?? y}
with a nullish coalescing operator \jscode{??}.  However, it is not true because
of the short-circuiting for assignments:
\begin{lstlisting}[style=JS]
       const x = 0;  x ??= 1;    // x === 0
       const y = 0;  y = y ?? 1; // TypeError
\end{lstlisting}
and the named evaluation for anonymous functions:
\begin{lstlisting}[style=JS]
   let f;  f ??= () => {};      // f.name === "f"
   let g;  g = g ?? (() => {}); // g.name === ""
\end{lstlisting}
Such misunderstanding of complex semantics in various language features causes
the wrong implementation of JavaScript applications or even JavaScript engines.

Thus, researchers have presented several tools for a better understanding of
JavaScript semantics. For example, JSExplain~\cite{jsexplain} is a reference
interpreter that provides step-by-step execution of a given JavaScript program
by closely following the English sentences of the specification. Several
JavaScript language designers accept the concept of reference interpreters and
implement Narcissus~\cite{narcissus} and engine262~\cite{engine262} to use in
the language design process. On the other hand, TC39 feels the need for type
information in ECMAScript to enhance its readability and has started internal
discussions on manual type annotations for each abstract
algorithm\footnote{https://github.com/tc39/ecmarkup/issues/173}. To meet the
demand of type information, \citet{jstar} presents a tool named $\jstar$ to
perform a type analysis for a given ECMAScript. Besides, they detect
type-related specification bugs in ES12 using the type analysis result. Finally,
$\jest$~\cite{jest} is another tool that synthesizes JavaScript conformance
tests from a given ECMAScript. Thus, it allows users to understand language
semantics using test programs instead of English sentences in the specification.

However, existing tools have two limitations: 1) manual update for evolving
language specification and 2) no user interactions to select target semantics.
Among existing tools, JavaScript reference interpreters require manual updates
when ECMAScript evolves. Until 2015, it was not a critical problem because
ECMAScript had rarely evolved. However, TC39 decided to annually release the
specification with a massive update in ECMAScript 6 (ES6, 2015)~\cite{es6} in
2015. Moreover, they published the specification as an open-source project in a
GitHub repository to quickly adapt users' demands to the language. For example,
JSExplain also requires manual updates for new specifications. Thus, it still
supports only ECMAScript 5.1 (ES5.1, 2011)~\cite{es5}, while already seven more
versions from ES6 to ES12 have been released. Another limitation is that
existing tools do not support user interactions to select target semantics to
get additional information. For example, $\jstar$ and $\jest$ do not interact
with users to give more information helpful for semantics understanding.
Instead, they only extract type information and conformance tests for a given
ECMAScript. On the other hand, while JSExplain supports step-by-step execution
as user interaction with several debugger-like features, it can deal with only
the execution trace of a single JavaScript program.

In this paper, we present $\tool$, a \textbf{J}ava\textbf{S}cript
\textbf{I}nteractive \textbf{S}pecification. It is the first reference
interpreter automatically synchronized with a given ECMAScript supporting
step-by-step program execution. Fortunately, \citet{jiset} presents a tool named
$\jiset$ that extracts a mechanized specification from a given specification.
The extracted mechanized specification is written in a specification language
$\ires$ and executable with a JavaScript program. Thus, we implement our tool
using $\jiset$ and design debugger-like features: stepping controls, state
visualization, and breakpoints with algorithm names or JavaScript program
points. Moreover, our tool supports two different types of user interactions to
select target semantics: \textit{syntactic view selection} and \textit{algorithm
step selection}. First, we introduce a way to reduce abstract algorithms using a
partial evaluation~\cite{peval, peval-survey, trans-ai} with a \textit{syntactic
view}. A syntactic view is a JavaScript Abstract Syntax Tree (AST) defined with
abstracted nodes. We formally define a partial evaluation with syntactic views
and prove its semantics preservation under the given syntactic view. Second, we
also present a way to extract a small-sized program for each algorithm step
using delta debugging~\cite{delta-debugging} with a given set of programs. For
the f first step, we execute each program using $\tool$ to compute semantics
coverage in ECMAScript and filter out unnecessary programs. Then, we perform
delta debugging for each program to reduce program sizes without loss of
semantics coverage.

Our contributions are as follows:
\begin{itemize}
  \item We present $\tool$, a JavaScript interactive specification. It is the
    first reference interpreter automatically synchronized with a given
    ECMAScript supporting step-by-step program execution.

  \item We introduce two different types of user interactions to select target
    semantics and implement them in $\tool$: 1) a \textit{syntactic view
    selection} to reduce abstract algorithms using \textit{partial evaluation}
    and 2) an \textit{algorithm step selection} to extract a small-sized
    JavaScript program related to the selected step using \textit{delta
    debugging} with a given set of programs.

  \item We evaluate $\tool$ with the latest ECMAScript, ES12, and experimentally
    show the effectiveness of two different target semantics selections with
    \inred{177} syntactic views and \inred{19,839} programs. Our tool
    successfully supports step-by-step program execution in ES12. It reduces
    \inred{18.9}\% algorithm steps for each syntactic view and provides
    \inred{2.3} lines (\inred{152.3} bytes) of a program for each step on
    average.
\end{itemize}